---
title: "Analysis of cant estimates"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(patchwork)
library(scico)
library(scales)
library(metR)
library(marelac)
library(kableExtra)
library(threejs)
```


```{r read_parameters, include = FALSE}
parameters <-
  read_rds(here::here("data",
                       "parameters.rds"))

slab_breaks <- c(parameters$slabs_Atl[1:12],Inf)
```

```{r read_mask_files, include = FALSE}
basinmask <-
  read_csv(
    here::here(
      "data/World_Ocean_Atlas_2018/_summarized_files",
      "basin_mask_WOA18.csv"
    )
  )
basinmask_AIP <-
  read_csv(
    here::here(
      "data/World_Ocean_Atlas_2018/_summarized_files",
      "basin_mask_WOA18_AIP.csv"
    )
  )
landmask <-
  read_csv(
    here::here(
      "data/World_Ocean_Atlas_2018/_summarized_files",
      "land_mask_WOA18.csv"
    )
  )

section_global_coordinates <-
  read_csv(here::here("data",
                       "section_global_coordinates.csv"))
```

```{r read_functions, include = FALSE}
source(here::here("code", "plotting_functions.R"))
```

```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

# Data sources

cant estimates from this study:

- Mean and SD per grid cell (lat, lon, depth)
- Zonal mean and SD (basin, lat, depth)
- Inventories (lat, lon)


```{r read_cant_files}


cant_3d <-
  read_csv(here::here("data/mapping/_summarized_files",
                         "cant_3d.csv"))

cant_zonal <-
  read_csv(here::here("data/mapping/_summarized_files",
                         "cant_zonal.csv"))

cant_predictor_average_zonal <-
  read_csv(here::here("data/mapping/_summarized_files",
                         "cant_predictor_zonal.csv"))

cant_inventory <-
  read_csv(here::here("data/mapping/_summarized_files",
                         "cant_inv.csv"))

```

C* estimates from this study:

- Mean and SD per grid cell (lat, lon, depth)
- Zonal mean and SD (basin, lat, depth)

```{r read_cstar_files}


Cstar_average <-
  read_csv(here::here("data/mapping/_summarized_files",
                         "Cstar_average.csv"))

Cstar_average_zonal <-
  read_csv(here::here("data/mapping/_summarized_files",
                         "Cstar_average_zonal.csv"))


```

Cleaned GLODAPv2_2020 file as used in this study

```{r read_GLODAP_used_for_fitting}

GLODAP <-
  read_csv(
    here::here(
      "data/GLODAPv2_2020/_summarized_data_files",
      "GLODAP_MLR_fitting_ready.csv"
    )
  )

```

# Color scale

For ease of comparison with Gruber et al (2019) we adapt their color scale, including the ranges and breaks applied in various types of visualizations.

```{r define_Gruber_rainbow}

rgb2hex <- function(r, g, b)
  rgb(r, g, b, maxColorValue = 100)

cols = c(rgb2hex(95, 95, 95),
         rgb2hex(0, 0, 95),
         rgb2hex(100, 0, 0),
         rgb2hex(100, 100, 0))

Gruber_rainbow <- colorRampPalette(cols)

rm(rgb2hex, cols)

```

# cant budgets

Global cant inventories were estimated in Pg-C. Please note that here we only added positive cant values in the upper `r parameters$inventory_depth`m and do not apply additional corrections for areas not covered.

```{r calculate_global_inventory}

cant_inventory <- left_join(cant_inventory, basinmask_AIP) 

cant_inventory_budget <- cant_inventory %>% 
  mutate(surface_area = earth_surf(lat, lon),
         cant_inv_grid = cant_inv*surface_area) %>% 
  group_by(eras, basin_AIP) %>% 
  summarise(cant_total = sum(cant_inv_grid)*12*1e-15,
            cant_total = round(cant_total,1)) %>% 
  ungroup() %>% 
  arrange(desc(eras)) %>% 
  pivot_wider(values_from = cant_total, names_from = basin_AIP) %>% 
  mutate(total = Atlantic + Indian + Pacific)

cant_inventory_budget %>% 
  kableExtra::kable() %>% 
  add_header_above() %>%
  kable_styling(full_width = FALSE)

rm(cant_inventory_budget)

```



# cant - positive

In a first series of plots we explore the distribution of cant, taking only positive estimates into account (positive here refers to the mean cant estimate across 10 eMLR model predictions available for each grid cell). Negative values were set to zero before calculating mean sections and inventories.

## Zonal mean sections

```{r cant_pos_zonal_mean_sections}

cant_zonal <- cant_zonal %>% 
  mutate(eras = factor(eras, c("JGOFS_GO", "GO_new")))

breaks <- c(seq(0,18,1),Inf)
breaks_n <- length(breaks) - 1

cant_zonal <- cant_zonal %>% 
  mutate(cant_pos_mean_int = cut(cant_pos_mean, 
                                breaks,
                                right = FALSE))


zonal_section <- function(df, i_basin_AIP, i_eras, var) {
  
  name_var <- var
  var <- sym(var)
  
  lat_max <- max(df$lat)
  lat_min <- min(df$lat)
  
  df_sub <- df %>% 
    filter(eras == i_eras,
           basin_AIP == i_basin_AIP)
  
  surface <- df_sub %>%
    ggplot(aes(lat, depth, z = !!var)) +
    geom_contour_filled(breaks = breaks) +
    scale_fill_manual(values = Gruber_rainbow(breaks_n),
                      name = "cant") +
    coord_cartesian(expand = 0,
                    ylim = c(500, 0),
                    xlim = c(lat_min, lat_max)) +
    scale_y_reverse() +
    scale_x_continuous(breaks = seq(-100,100,20)) +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    ) +
    labs(y = "Depth (m)",
         title = paste("Basin:", i_basin_AIP, "| eras:", i_eras))
  
  deep <- df_sub %>%
    ggplot(aes(lat, depth, z = !!var)) +
    geom_contour_filled(breaks = breaks) +
    scale_fill_manual(values = Gruber_rainbow(breaks_n),
                      name = "cant") +
    scale_y_reverse() +
    scale_x_continuous(breaks = seq(-100,100,20)) +
    coord_cartesian(expand = 0,
                    ylim = c(3000, 500),
                    xlim = c(lat_min, lat_max)) +
    labs(x = "latitude (°N)", y = "Depth (m)")
  
  surface / deep +
    plot_layout(guides = "collect")
  
}


for (i_basin_AIP in unique(cant_zonal$basin_AIP)) {
  for (i_eras in rev(unique(cant_zonal$eras))) {
    
    print(zonal_section(cant_zonal,
                        i_basin_AIP = i_basin_AIP,
                        i_eras = i_eras,
                        "cant_pos_mean"))
    
  }
  
}

rm(breaks, breaks_n)
  
```

## Isoneutral slab distribution

Mean of positive cant within each horizontal grid cell (lon x lat) per isoneutral slab.

Please note that:

- density slabs covering values >28.1 occur by definition only either in the Atlantic or Indo-Pacific basin
- gaps in the maps represent areas where (thin) density layers fit between discrete depth levels used for mapping


```{r cant_pos_gamma_slab_maps, fig.asp=2.5}

cant_gamma_maps <- cant_3d %>% 
  group_by(lat, lon, gamma_slab, eras) %>% 
  summarise(cant_pos = mean(cant_pos, na.rm = TRUE)) %>% 
  ungroup()


breaks <- c(seq(0,16,2),Inf)
breaks_n <- length(breaks) - 1

cant_gamma_maps <- cant_gamma_maps %>% 
  mutate(cant_pos_int = cut(cant_pos, 
                                breaks,
                                right = FALSE)) %>% 
  mutate(eras = factor(eras, c("JGOFS_GO", "GO_new")))


ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat),
              fill = "grey30") +
  geom_raster(data = cant_gamma_maps,
              aes(lon, lat, fill = cant_pos_int)) +
  scale_fill_manual(values = Gruber_rainbow(breaks_n)) +
  facet_grid(gamma_slab ~ eras) +
  coord_quickmap(expand = 0) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "top")

rm(cant_gamma_maps, breaks, breaks_n)
```

## Inventory map

Column inventory of positive cant between the surface and `r parameters$inventory_depth`m water depth per horizontal grid cell (lat x lon).

```{r cant_pos_inventory_map, fig.asp=1}

breaks <- c(seq(0,16,2),Inf)
breaks_n <- length(breaks) - 1

cant_inventory <- cant_inventory %>% 
  mutate(cant_inv_pos_int = cut(cant_inv_pos, 
                                breaks,
                                right = FALSE)) %>% 
  mutate(eras = factor(eras, c("JGOFS_GO", "GO_new")))

cant_inventory %>%
  ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat), fill = "grey30") +
  geom_raster(aes(lon, lat, fill = cant_inv_pos_int)) +
  coord_quickmap(expand = 0) +
  scale_fill_manual(values = Gruber_rainbow(breaks_n)) +
  # scale_fill_scico_d(palette = "batlow", direction = -1) +
  facet_wrap( ~ eras, ncol = 1) +
  theme(axis.title = element_blank())

```


## Global section

### JGOFS_GO

```{r cant_sections_positive_mean_one_lon_JGOFS_GO}

section_global(cant_3d %>% filter(eras == "JGOFS_GO"),
                    "cant_pos")

```

### GO_new

```{r cant_sections_positive_mean_one_lon_GO_new}

section_global(cant_3d %>% filter(eras == "GO_new"),
                    "cant_pos")

```


# cant - all

In a second series of plots we explore the distribution of cant, taking positive and negative estimates into account.

## Zonal mean sections

```{r cant_all_zonal_mean_sections}
section_zonal_average_divergent(cant_zonal,
                                "cant_mean",
                                "gamma_mean")
```


## Isoneutral slab distribution

Mean of cant within each horizontal grid cell (lon x lat) per isoneutral slab.

Please note that:

- density slabs covering values >28.1 occur by definition only either in the Atlantic or Indo-Pacific basin
- gaps in the maps represent areas where (thin) density layers fit between discrete depth levels used for mapping


```{r cant_all_gamma_slab_maps, fig.asp=2.5}

cant_gamma_maps <- cant_3d %>%
  group_by(lat, lon, gamma_slab, eras) %>%
  summarise(cant = mean(cant, na.rm = TRUE)) %>%
  ungroup()


breaks <- c(-Inf, seq(-16, 16, 4), Inf)
breaks_n <- length(breaks) - 1

cant_gamma_maps <- cant_gamma_maps %>%
  mutate(cant_int = cut(cant,
                        breaks,
                        right = FALSE)) %>% 
  mutate(eras = factor(eras, c("JGOFS_GO", "GO_new")))

ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat),
              fill = "grey30") +
  geom_raster(data = cant_gamma_maps,
              aes(lon, lat, fill = cant_int)) +
  scale_fill_brewer(palette = "RdBu",
                    direction = -1) +
  facet_grid(gamma_slab ~ eras) +
  coord_quickmap(expand = 0) +
  theme(
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.position = "top"
  )

rm(cant_gamma_maps, breaks, breaks_n)
```

## Inventory map

Column inventory of positive cant between the surface and `r parameters$inventory_depth`m water depth per horizontal grid cell (lat x lon).

```{r cant_all_inventory_map, fig.asp=1}

breaks <- c(-Inf, seq(-16,16,4),Inf)
breaks_n <- length(breaks) - 1

cant_inventory <- cant_inventory %>% 
  mutate(cant_inv_int = cut(cant_inv, 
                                breaks,
                                right = FALSE))
cant_inventory %>% 
  ggplot() +
    geom_raster(data = landmask,
                aes(lon, lat), fill = "grey80") +
    geom_raster(aes(lon, lat, fill = cant_inv_int)) +
    coord_quickmap(expand = 0) +
  scale_fill_brewer(palette = "RdBu",
                    direction = -1) +
  facet_wrap(~eras, ncol = 1) +
    theme(
      axis.title = element_blank()
    )


rm(breaks, breaks_n)

```

# cant - standard deviation 

## Across models

Standard deviation across cant from all MLR models was calculate for each grid cell (XYZ). The zonal mean of this standard deviation should reflect the uncertainty associated to the predictor selection within each slab and era. 

```{r cant_sections_sd_models}

section_zonal_average_continous(cant_zonal,
                                "cant_sd_mean",
                                "gamma_mean")

```

## Across basins

Standard deviation of mean cant values was calculate across all longitudes. This standard deviation should reflect the zonal variability of cant within the basin and era.

```{r cant_sections_sd_cant}

section_zonal_average_continous(cant_zonal,
                                "cant_sd",
                                "gamma_mean")

```

## Correlation

### cant vs model SD

```{r cant_vs_sd_by_basin_era}

cant_3d <- cant_3d  %>% 
  mutate(eras = factor(eras, c("JGOFS_GO", "GO_new")))

cant_3d %>% 
  ggplot(aes(cant, cant_sd)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 10) +
  geom_bin2d() +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10",
                       name = "log10(n)") +
  facet_grid(basin_AIP ~ eras)

```

```{r cant_vs_sd_by_basin_gamma, fig.asp=2}

cant_3d %>% 
  ggplot(aes(cant, cant_sd)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 10) +
  geom_bin2d() +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10",
                       name = "log10(n)") +
  facet_grid(gamma_slab ~ basin_AIP)

```

### cant vs regional SD

```{r cant_vs_sd_by_basin_era_zonal}

cant_zonal %>% 
  ggplot(aes(cant_mean, cant_sd)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 10) +
  geom_bin2d() +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10",
                       name = "log10(n)") +
  facet_grid(basin_AIP ~ eras)

```

```{r cant_vs_sd_by_basin_gamma_zonal, fig.asp=2}

cant_zonal %>% 
  ggplot(aes(cant_mean, cant_sd)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 10) +
  geom_bin2d() +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10",
                       name = "log10(n)") +
  facet_grid(gamma_slab ~ basin_AIP)

```

# cant - predictor contribution

```{r cant_section_predictor_contribution}

cant_predictor_average_zonal <- cant_predictor_average_zonal %>% 
  mutate(eras = factor(eras, c("JGOFS_GO", "GO_new")))


for (variable in c(
  "cant_intercept",
  "cant_aou",
  "cant_oxygen",
  "cant_phosphate",
  "cant_phosphate_star",
  "cant_silicate",
  "cant_tem",
  "cant_sal")) {

print(
section_zonal_average_continous(cant_predictor_average_zonal,
                                variable,
                                "gamma")
)
    
}


rm(variable)
```


# Neutral density

## Slab depth

Please note that:

- density slabs covering values >28.1 occur by definition only either in the Atlantic or Indo-Pacific basin
- predictor density slabs are only shown for the upper `r parameters$inventory_depth`m as used for the mapping, whereas GLODAP observations are displayed for the entire water column as used for fitting eMLRs (in both cases shallow waters are excluded at low density)

```{r gamma_maps, fig.asp=2.5}

gamma_maps <- cant_3d %>% 
  group_by(lat, lon, gamma_slab) %>% 
  summarise(depth_max = max(depth, na.rm = TRUE)) %>% 
  ungroup()

GLODAP_obs_coverage <- GLODAP %>% 
  count(lat, lon, gamma_slab, era) %>% 
  mutate(era = factor(era, c("JGOFS_WOCE", "GO_SHIP", "new_era")))
  
ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat),
              fill = "grey80") +
  geom_raster(data = gamma_maps,
              aes(lon, lat, fill = depth_max)) +
  geom_raster(data = GLODAP_obs_coverage,
              aes(lon, lat), fill = "red") +
  facet_grid(gamma_slab ~ era) +
  coord_quickmap(expand = 0) +
  scale_fill_viridis_c(direction = -1) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "top")

rm(gamma_maps, GLODAP_obs_coverage)
```

## Surface maps

```{r gamma_surface_map}

map_climatology_discrete(cant_3d, "gamma_slab")

```


## Zonal sections

The mean zonal distribution of neutral densities was calculated. **CAVEAT:** Due to practical reasons, binning here does not include the two highest isoneutral density slabs in the Atlantic, yet.

### Mean

```{r gamma_sections_mean}

cant_zonal %>% 
  filter(eras == "JGOFS_GO") %>% 
  ggplot(aes(lat, depth, z = gamma_mean)) +
  geom_contour_filled(breaks = slab_breaks) +
  geom_contour(breaks = slab_breaks,
               col = "white") +
  geom_text_contour(breaks = slab_breaks,
               col = "white",
               skip = 1) +
  scale_fill_viridis_d(name = "Gamma",
                       direction = -1) +
  scale_y_reverse() +
  scale_x_continuous(breaks = seq(-100, 100, 20)) +
  coord_cartesian(expand = 0) +
  guides(fill = guide_colorsteps(barheight = unit(10, "cm"))) +
  facet_grid(basin_AIP~.)

```

### SD

Higher SD of gamma in shallow, subtropical waters results from a more pronounced longitudinal variability.

```{r gamma_sections_sd}

cant_zonal %>% 
  filter(eras == "JGOFS_GO") %>% 
  ggplot(aes(lat, depth, z = gamma_sd)) +
  geom_contour_filled() +
  scale_fill_viridis_d(name = "Gamma SD",
                       direction = -1) +
  scale_y_reverse() +
  scale_x_continuous(breaks = seq(-100, 100, 20)) +
  coord_cartesian(expand = 0) +
  guides(fill = guide_colorsteps(barheight = unit(10, "cm"))) +
  facet_grid(basin_AIP~.)

```


# Cstar

## Zonal mean sections

```{r Cstar_zonal_mean_era}

Cstar_average_zonal <- Cstar_average_zonal %>% 
    mutate(era = factor(era, c("JGOFS_WOCE", "GO_SHIP", "new_era"))) 

Cstar_average_zonal %>% 
    ggplot(aes(lat, depth, z = Cstar_mean_mean)) +
    geom_contour_filled(bins = 11) +
    scale_fill_viridis_d(name = "Cstar") +
    geom_contour(aes(lat, depth, z = gamma_mean_mean),
                 breaks = slab_breaks,
                 col = "white") +
    geom_text_contour(
      aes(lat, depth, z = gamma_mean_mean),
      breaks = slab_breaks,
      col = "white",
      skip = 1
    ) +
    scale_y_reverse() +
    scale_x_continuous(breaks = seq(-100,100,20)) +
    coord_cartesian(expand = 0) +
    guides(fill = guide_colorsteps(barheight = unit(10, "cm"))) +
    facet_grid(basin_AIP ~ era)



```

# Sections by model

Zonal sections plots are produced for every 20° longitude, each era and for all models individually and can be downloaded [here](https://github.com/jens-daniel-mueller/cant_eMLR/tree/master/output/figure/mapping){target="_blank"}.


# Known issues

Deviations between this study and the results by Gruber et al (2019), short G19, for the same period, might be attributable to following known differences in the implementation of the eMLR(C*) method:

- GLODAPv2_2020 here vs an extended version of GLODAPv2 in G19
- flagging: Here, we accept f flags 0 and 2 (except for tco2, where only 0 is accepted). G19 claim to use 0 throughout, yet have a high coverage of talk observations in the SE Pacific  
- Neutral density calculation: Here and in GLODAPv2_2020 a polynomial approximation is used, whereas G19 uses the original Matlab code  
- Predictor climatology: Here we used WOA18, whereas G19 used WOA13
- Missing data in the GLODAP mapped climatology, eg NO3 at surface, where not filled in this study
- cant on neutral density levels calculate as slab mean, rather than on one surface
- Here, surface delta cant were calculated based on Luecker constants, rather than Mehrbach as in G19
- Here, pCO2 was calculated from DIC/TA Climatology
