---
title: "Nomenclature"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r library, message=FALSE, warning=FALSE}

```


# Variables

Variables from source data files are converted and harmonized to satisfy following naming convention throughout the project:

- coordinates on 1x1 degree grid
  - lon: longitude (20.5 to 379.5 °E)
  - lat: latitude (-89.5 to 89.5 °N)
- depth: water depth (m)
- bottomdepth: bottomdepth (m)

- sal: salinity (Check scales!)
- tem: insitu temperature in deg C (Check scales!)
- theta: potential temperature in deg C (Check scales!)
- gamma: neutral density

- cant: anthropogenic CO~2~ (mol kg^-1^)

# Data sets / objects

# Chunk label naming within .Rmd files

- read_xxx: open new data set
- clean_xxx: subset rows of a data set
- calculate_xxx: perform calculations on a data set (add or modify rows)
- write_xxx: write summarized data file to disk

# Folder structure

- data: contains all data, not synced to Github
  - subfolder for each data product
    - _summarized_data_files: data sets created along analysis


# Open tasks

- check temperature and salinity scales 
