---
title: "Nomenclature"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r library, message=FALSE, warning=FALSE}

```


# Variables

Variables from source data files are converted and harmonized to following variables:

- 1x1 degree coordinate grid
  - lon: 20.5 to 379.5°E longitude (in °E)
  - lat: -89.5 to 89.5°N latitude (in °N)
- depth: water depth in m
- bottomdepth: bottomdepth in m

- sal: salinity (Check scales!)
- tem: insitu temperature in deg C (Check scales!)
- theta: potential temperature in deg C (Check scales!)
- gamma: neutral density

- Cant: anthropogenic CO~2~ in umol kg-1

# Data sets / objects

# Chunk labels

- read_xxx: open new data set
- clean_xxx: subset rows of a data set
- write_xxx: write summarized data file to disk

# Folder structure

- data: contains all data, not synced to Github
  - subfolder for each data product
    - _summarized_data_files: data sets created along analysis


# Open tasks

- check temperature and salinity scales 
