---
title: "Data base"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(patchwork)
```


```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

# Read GLODAP master file

- Data source: `GLODAPv2.2020_Merged_Master_File.csv` from [glodap.info](https://www.glodap.info/)

```{r read_GLODAPv2_2020_merged_master_file, eval=FALSE}

GLODAP <- read_csv(here::here("data", "GLODAPv2.2020_Merged_Master_File.csv"),
                   na = "-9999",
                   col_types = cols(.default = col_double()))

# relevant columns
GLODAP <- GLODAP %>% 
  select(cruise:phtsqc)

GLODAP <- GLODAP %>% 
  mutate(date = ymd(paste(year, month, day)),
         decade = as.factor(floor(year / 10) * 10)) %>% 
  relocate(date, decade)


```

# Clean data

- subsetted rows with 
  - paired tco2, talk, and (!) phosphate observations
  - flag f: 2
  - flag qc: 1

```{r clean_GLODAP_master_file,eval=FALSE}

# select rows with tco2, good flag
GLODAP_clean <- GLODAP %>% 
  filter(!is.na(tco2))

GLODAP_clean <- GLODAP_clean %>% 
  filter(tco2f == "2")

GLODAP_clean <- GLODAP_clean %>% 
  filter(tco2qc == "1")


# select rows with talk, good flag

GLODAP_clean <- GLODAP_clean %>% 
  filter(!is.na(talk))

GLODAP_clean <- GLODAP_clean %>% 
  filter(talkf == "2")

GLODAP_clean <- GLODAP_clean %>% 
  filter(talkqc == "1")


# select rows with talk, good flag

GLODAP_clean <- GLODAP_clean %>% 
  filter(!is.na(phosphate))

GLODAP_clean <- GLODAP_clean %>% 
  filter(phosphatef == "2")

GLODAP_clean <- GLODAP_clean %>% 
  filter(phosphateqc == "1")


# date and decade column
GLODAP_clean  %>%  write_csv(here::here("data/_summarized_data_files", "GLODAPv2.2020_Merged_Master_File_clean.csv"))

```




# Overview GLODAPv2-2020

## Nr of observations

```{r calculate_stats, eval=FALSE}

GLODAP_stats <- GLODAP %>% 
  summarise(total = n(),
            tco2 = sum(!is.na(tco2)),
            tco2_flag = sum(!is.na(tco2) & tco2f == 2 & tco2qc == 1),
            talk = sum(!is.na(talk)),
            talk_f_qc = sum(!is.na(talk) & talkf == 2 & talkqc == 1),
            phosphate = sum(!is.na(phosphate)),
            phosphate_f_qc = sum(!is.na(phosphate) & phosphatef == 2 & phosphateqc == 1))

GLODAP_stats_2 <- GLODAP %>% 
  summarise(tco2_talk = sum(!is.na(tco2) & !is.na(talk)),
            tco2_phosphate = sum(!is.na(tco2) & !is.na(phosphate)),
            tco2_talk_phosphate = sum(!is.na(tco2) & !is.na(talk) & !is.na(phosphate)),
            tco2_talk_phosphate_f_qc = sum(!is.na(tco2) & !is.na(talk) & !is.na(phosphate) &
                                        tco2f == 2 & tco2qc == 1 &
                                        talkf == 2 & talkqc == 1 &
                                        phosphatef == 2 & phosphateqc == 1),
           incl_pH = sum(!is.na(tco2) & !is.na(talk) & !is.na(phosphate) &
                                        tco2f == 2 & tco2qc == 1 &
                                        talkf == 2 & talkqc == 1 &
                                        phosphatef == 2 & phosphateqc == 1 &
                                        !is.na(phts25p0)))

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_2)

rm(GLODAP_stats_2)

GLODAP_stats_long <- GLODAP_stats %>% 
  pivot_longer(1:12, names_to = "parameter", values_to = "n")

GLODAP_stats_long  %>%  write_csv(here::here("data/_summarized_data_files",
                                             "GLODAPv2.2020_stats.csv"))

```

```{r GLODAP_stats_barplot}

GLODAP_stats_long <-  read_csv(here::here("data/_summarized_data_files",
                                             "GLODAPv2.2020_stats.csv"))

GLODAP_stats_long %>% 
  ggplot(aes(parameter, n))+
  geom_col()+
  coord_flip()

rm(GLODAP_stats_long)

```

## Data coverage

Spatio-temporal distribution of non-flagged, paired `talk`, `tco2`, and `phosphate` data.

```{r reopen_GLODAP}

GLODAP <- read_csv(here::here("data/_summarized_data_files",
                              "GLODAPv2.2020_Merged_Master_File_clean.csv"),
                   guess_max = 1e5)

GLODAP <- GLODAP %>% 
  mutate(decade = as.factor(decade))

```

## Gridding

```{r spatial_gridding}

GLODAP <- GLODAP %>% 
  mutate(lat_grid = cut(latitude, seq(-90, 90, 5), seq(-87.5, 87.5, 5)),
         lat_grid = as.numeric(as.character(lat_grid)),
         lon_grid = cut(longitude, seq(-180, 180, 5), seq(-177.5, 177.5, 5)),
         lon_grid = as.numeric(as.character(lon_grid)))

GLODAP_map_year <- GLODAP %>% 
  group_by(year, lat_grid, lon_grid) %>% 
  tally() %>% 
  ungroup()

GLODAP_map_decade <- GLODAP %>% 
  group_by(decade, lat_grid, lon_grid) %>% 
  tally() %>% 
  ungroup()

GLODAP_hovmoeller_year <- GLODAP %>% 
  group_by(year, lat_grid) %>% 
  tally() %>% 
  ungroup()


```

## Histograms

```{r histogramm_observations}

GLODAP %>% 
  ggplot(aes(latitude, fill=decade))+
  geom_histogram(binwidth = 10, col="black", boundary = 0)+
  scale_x_continuous(breaks = seq(-85,90,10))+
  scale_fill_viridis_d()+
  coord_flip(expand = 0)+
  labs(title = "paired tco2 + talk + phosphate observations")

```

## Hovmoeller (Lat-time)

```{r coverage_Hovmoeller}

GLODAP_hovmoeller_year %>% 
  ggplot(aes(year, lat_grid, fill=log10(n)))+
  geom_tile()+
  scale_fill_viridis_c(option = "magma", direction = -1)

rm(GLODAP_hovmoeller_year)

```

## Maps

```{r coverage_maps_decade, fig.asp=2.5}

mapWorld <- borders("world", colour="gray60", fill="gray60")

GLODAP_map_decade %>%
  ggplot(aes(lon_grid, lat_grid, fill=log10(n)))+
  mapWorld+
  geom_raster()+
  scale_fill_viridis_c(option = "magma", direction = -1)+
  scale_x_continuous(breaks = seq(-180, 180, 30))+
  scale_y_continuous(breaks = seq(-90, 90, 30))+
  coord_quickmap(expand = FALSE)+
  facet_wrap(~decade, ncol=1)

rm(GLODAP_map_decade)

```

```{r coverage_maps_year, fig.asp=3.5}

GLODAP_map_year %>%
  filter(year > 1990) %>% 
  ggplot(aes(lon_grid, lat_grid, fill=log10(n)))+
  mapWorld+
  geom_raster()+
  scale_fill_viridis_c(option = "magma", direction = -1)+
  coord_quickmap(expand = FALSE)+
  facet_wrap(~year, ncol = 2)+
  theme(axis.title = element_blank(),
        axis.text = element_blank())

rm(GLODAP_map_year)

```

## Individual cruise sections

```{r individual_cruises_clean_triplicates, eval = FALSE}

cruises <- GLODAP %>% 
  group_by(cruise) %>% 
  summarise(date_mean = mean(date, na.rm = TRUE),
            n = n()) %>% 
  ungroup() %>% 
  arrange(-n, cruise)

GLODAP <- full_join(GLODAP, cruises)

mapWorld <- borders("world", colour="gray60", fill="gray60")

n <- 0
for (i_cruise in unique(cruises$cruise)) {

#i_cruise <- unique(cruises$cruise)[1]
n <- n+1
print(n)  
  
GLODAP_cruise <- GLODAP %>%
  filter(cruise == i_cruise) %>% 
  arrange(year, month, day, hour, minute)

cruises_cruise <- cruises %>%
  filter(cruise == i_cruise)
  
map <- GLODAP_cruise %>%
  ggplot(aes(longitude, latitude))+
  mapWorld+
  geom_point(aes(col=date))+
  geom_path()+
  coord_quickmap(expand = FALSE)+
  scale_color_viridis_c(trans = "date")+
  labs(title = paste("Mean date:", cruises_cruise$date_mean,
                     "| cruise:", cruises_cruise$cruise,
                     "| n(samples):", cruises_cruise$n))


lon_section <- GLODAP_cruise %>%
  ggplot(aes(longitude, depth))+
  scale_y_reverse()+
  scale_color_viridis_c()

lon_tco2 <- lon_section+
  geom_point(aes(col=tco2))

lon_talk <- lon_section+
  geom_point(aes(col=talk))

lon_phosphate <- lon_section+
  geom_point(aes(col=phosphate))


lat_section <- GLODAP_cruise %>%
  ggplot(aes(latitude, depth))+
  scale_y_reverse()+
  scale_color_viridis_c()

lat_tco2 <- lat_section+
  geom_point(aes(col=tco2))

lat_talk <- lat_section+
  geom_point(aes(col=talk))

lat_phosphate <- lat_section+
  geom_point(aes(col=phosphate))

map /
((lat_tco2 / lat_talk / lat_phosphate) |
(lon_tco2 / lon_talk / lon_phosphate))

ggsave(here::here("output/plots/data/all_cruises_clean_triplicates",
                  paste("GLODAP_cruise_",
                        cruises_cruise$n,
                        cruises_cruise$date_mean,
                        cruises_cruise$cruise,
                        ".png",
                        sep = "_")),
                  width = 9, height = 9)

rm(map,
   lon_section, lat_section,
   lat_tco2, lat_talk, lat_phosphate, lon_tco2, lon_talk, lon_phosphate,
   GLODAP_cruise, cruises_cruise)

}


# library("rnaturalearth")
# library("rnaturalearthdata")
# 
# world <- ne_countries(scale = "small", returnclass = "sf")
# class(world)
# 
# ggplot(data = world) +
#     geom_sf()
# 
# ggplot()+
#   mapWorld

```
