---
title: "Data base"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(patchwork)
```


```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

# Open required data sets

## GLODAP

```{r reopen_GLODAP}

GLODAP <- read_csv(here::here("data/GLODAPv2_2020/_summarized_data_files",
                              "GLODAPv2.2020_clean.csv"))

```

## GLODAP Cant climatology

```{r open_GLODAP_Cant_climatology}

Cant_clim <- read_csv(here::here("data/GLODAPv2_2016b_MappedClimatologies/_summarized_files",
                                 "Cant.csv"))
```



# C*

## stoichiometric ratios

```{r set_stoichiometric_ratios}

rCP <- 117
rNP <- 16

```

The stoichiometric nutrient ratios for the production and mineralization of organic matter were set to:

- C/P: `r rCP`
- N/P: `r rNP`

## Calculation

```{r calculate_C_star}

GLODAP <- GLODAP %>% 
  mutate(rCP_phosphate = - rCP * phosphate,
         talk_05 = - 0.5 * talk,
         rNP_phosphate_05 = - 0.5 * rNP * phosphate,
         C_star = tco2 + rCP_phosphate + talk_05 + rNP_phosphate_05)

```

# Selected section plots

```{r select_cruises}

cruises_meridional <- c("1041")

# cruises_meridional <- c("1041","1042", "260",
#                         "2011", "393", "1031", "394", "395",
#                         "1088", "983")

# cruises_zonal <- c()

GLODAP_cruise <- GLODAP %>% 
  filter(cruise %in% cruises_meridional)

```

```{r meridional_sections, fig.asp=3}

mapWorld <- borders("world", colour="gray60", fill="gray60")

map <- GLODAP_cruise %>%
  ggplot(aes(lon, lat))+
  mapWorld+
  geom_point(aes(col=date))+
  geom_path()+
  coord_quickmap(expand = FALSE)+
  scale_color_viridis_c(trans = "date")

lat_section <- GLODAP_cruise %>%
  ggplot(aes(lat, depth))+
  scale_y_reverse()+
  scale_color_viridis_c()

lat_tco2 <- lat_section+
  geom_point(aes(col=tco2))

lat_talk <- lat_section+
  geom_point(aes(col=talk))

lat_phosphate <- lat_section+
  geom_point(aes(col=phosphate))

lat_rCP_phosphate <- lat_section+
  geom_point(aes(col=rCP_phosphate))

lat_talk_05 <- lat_section+
  geom_point(aes(col=talk_05))

lat_rNP_phosphate_05 <- lat_section+
  geom_point(aes(col=rNP_phosphate_05))

lat_C_star <- lat_section+
  geom_point(aes(col=C_star))

map / lat_tco2 / lat_talk / lat_phosphate / lat_rCP_phosphate /lat_talk_05 /lat_rNP_phosphate_05 / lat_C_star

rm(map, lat_tco2, lat_talk, lat_phosphate, lat_C_star)

```



# Reference year adjustment

The scaling factor Cant_apriori at a given location and depth was estimated for each reference year from XXX.
Note that eq. 6 in Clement and Gruber (2018) misses pCO2 pre-industrial in the denominator.

## Merge data sets

```{r merge_Cstar_Cant_pCO2}

```

## Perform adjustment

```{r adjust_reference_year, eval=FALSE}

GLODAP <- GLODAP %>% 
  mutate(C_star_ref = C_star - 
           ( (pCO2_atm(year) - pCO2_atm(year_ref)) / 
               (pCO2_atm(year_ref) - pCO2_atm(year_pi) ) * Cant_ref)

```

# MLR fitting

```{r predictor_variables, eval=FALSE}

temperature
salinity
phosphate
silicate
phosphate_star = phosphate + (oxygen / 170) - 1.95
oxygen
aou


```

```{r estimate_MLR_models, eval=FALSE}

basins <- c("Atlantic", "Indo_Pacific")
slabs <- c("")

for (i_basin in basins) {
  for (i_slab in slabs) {
    
    
  }
  
}

```


