---
title: "Mapping cstar"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = here::here("analysis/child/setup.Rmd")}
```

# Predictor fields

Currently, we use combined predictor fields:

- WOA18: S, T, and derived variables
- GLODAP16: Oxygen, PO4, NO3, Silicate, and derived variables

```{r read_predictor_file}

predictors <- 
    read_csv(here::here("data/mapping",
                         "W18_st_G16_opsn.csv"))

```


# Load MLR models

```{r load_eMLR_models}

lm_all_wide_cstar <-
  read_csv(here::here("data/eMLR",
                       "lm_all_wide_cstar.csv"))

```

# Merge MLRs + climatologies

```{r merge_model_coeff_predictor_climatology}

lm_all_wide_cstar <- lm_all_wide_cstar %>% 
  mutate(model = str_remove(model, "cstar ~ "))
         
cstar <- full_join(predictors, lm_all_wide_cstar)

rm(predictors, lm_all_wide_cstar)

```


# Map Cstar

## Apply MLRs to predictor

```{r calculate_cstar}

cstar <- cstar %>% 
  mutate(cstar =
          `coeff_(Intercept)` +
           coeff_aou * aou +
           coeff_oxygen * oxygen +
           coeff_phosphate * phosphate +
           coeff_phosphate_star * phosphate_star +
           coeff_silicate * silicate +
           coeff_sal * sal + 
           coeff_tem * tem)

cstar <- cstar %>% 
  select(lon, lat, depth, era, basin, cstar, gamma)

```

```{r cstar_average}

cstar_average <- cstar %>%
  fselect(lon, lat, depth, era, basin, cstar, gamma) %>% 
  fgroup_by(lon, lat, depth, era, basin) %>% {
   add_vars(fgroup_vars(.,"unique"),
            fmean(., keep.group_vars = FALSE),
            fsd(., keep.group_vars = FALSE) %>% add_stub(pre = FALSE, "_sd"))
  }

rm(cstar)

cstar_average_Atl <- cstar_average %>% 
  filter(basin == "Atlantic") %>% 
  mutate(gamma_slab = cut(gamma, parameters$slabs_Atl))

cstar_average_Ind_Pac <- cstar_average %>% 
  filter(basin == "Indo-Pacific") %>% 
  mutate(gamma_slab = cut(gamma, parameters$slabs_Ind_Pac))

cstar_average <- bind_rows(cstar_average_Atl, cstar_average_Ind_Pac)

rm(cstar_average_Atl, cstar_average_Ind_Pac)

```

## Mean cstar sections

For each basin and era combination, the zonal mean cstar is calculated. Likewise, sd is calculated for the averaging of the mean basin fields.

```{r Calculate_cstar_mean_sections}


cstar_average <- left_join(cstar_average,
                           basinmask %>% select(-basin))


cstar_average_zonal <- cstar_average %>%
  fselect(lat, depth, era, basin, basin_AIP,
          cstar, gamma, cstar_sd, gamma_sd) %>% 
  fgroup_by(lat, depth, era, basin, basin_AIP) %>% {
   add_vars(fgroup_vars(.,"unique"),
            fmean(., keep.group_vars = FALSE) %>% add_stub(pre = FALSE, "_mean"),
            fsd(., keep.group_vars = FALSE) %>% add_stub(pre = FALSE, "_sd"))
  }


cstar_average_zonal_Atl <- cstar_average_zonal %>% 
  filter(basin == "Atlantic") %>% 
  mutate(gamma_slab = cut(gamma_mean, parameters$slabs_Atl))

cstar_average_zonal_Ind_Pac <- cstar_average_zonal %>% 
  filter(basin == "Indo-Pacific") %>% 
  mutate(gamma_slab = cut(gamma_mean, parameters$slabs_Ind_Pac))

cstar_average_zonal <- bind_rows(cstar_average_zonal_Atl, cstar_average_zonal_Ind_Pac)

rm(cstar_average_zonal_Atl, cstar_average_zonal_Ind_Pac)


```

# Write csv

```{r write_cstar_files}

cstar_average %>%
    write_csv(here::here("data/output",
                         "cstar_3d.csv"))

cstar_average_zonal %>%
    write_csv(here::here("data/output",
                         "cstar_zonal.csv"))

```
