---
title: "Mapping"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(metR)
library(marelac)

```

```{r read_parameters, include = FALSE}

parameters <-
  read_rds(here::here("data",
                       "parameters.rds"))
```

```{r read_mask_files}

basinmask <- read_csv(here::here("data/World_Ocean_Atlas_2018/_summarized_files",
                                 "basin_mask_WOA18.csv"))

landmask <- read_csv(here::here("data/World_Ocean_Atlas_2018/_summarized_files",
                                 "land_mask_WOA18.csv"))

```

```{r read_parameters_functions, include = FALSE}
source(here::here("code", "plotting_functions.R"))
```

```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

# Required data

All required data sets were subsetted spatially in the read-in section *Data base*.
Currently, following data sets are used for mapping:

## GLODAPv2_2016b_MappedClimatologies

Following variables are currently used:  

- Salinity
- Temperature
- Phosphate (+Phosphate*)
- Silicate
- Oxygen (+AOU)


```{r load_GLODAPv2_2016b_MappedClimatologies}

variables <-
  c("salinity", "temperature", "oxygen", "PO4", "silicate", "NO3")

for (i_variable in variables) {
  temp <- read_csv(
    here::here(
      "data/GLODAPv2_2016b_MappedClimatologies/_summarized_files",
      paste(i_variable, ".csv", sep = "")
    )
  )
  
  if (exists("GLODAP_predictors")) {
    GLODAP_predictors <- full_join(GLODAP_predictors, temp)
  }
  
  if (!exists("GLODAP_predictors")) {
    GLODAP_predictors <- temp
  }
}

rm(temp, i_variable, variables)

# removed na's attributable to slightly different coverage of predictor fields
GLODAP_predictors <- GLODAP_predictors %>%
  drop_na()

GLODAP_predictors <- GLODAP_predictors %>%
  rename(sal = salinity,
         tem = temperature)

```


## World Ocean Atlas 2018

- Neutral density
- Basin mask

```{r load_WOA18}

WOA18_predictors <-
  read_csv(
    here::here(
      "data/World_Ocean_Atlas_2018/_summarized_files",
      "WOA18_predictors.csv"
    )
  )

```

## WOA13

Neutral densities and the basin mask based on WOA13 and provided by Dominic Clement are currently not used.

```{r load_WOA13_gamma, eval=FALSE}

WOA13 <-
  read_csv(
    here::here(
      "data/World_Ocean_Atlas_2013_Clement/_summarized_files",
      "WOA13_mask_gamma.csv"
    )
  )

WOA13_gamma <- WOA13 %>%
  select(-mask)

rm(WOA13)

```


# Join WOA18 + GLODAP

WOA18 and GLODAP predictor climatologies are merged. Only horizontal grid cells with observations from both predictor fields are kept.

**CAVEAT**: Coverage of GLODAP climatologies differs slightly for parameters (some are NA in some regions)

```{r join_WOA18_GLODAP_predictors, fig.asp=0.5}

predictors <- full_join(
  GLODAP_predictors %>% select(-c(sal,tem)),
  WOA18_predictors)

predictors <- predictors %>% 
  drop_na()

# rm(GLODAP_predictors, WOA18_predictors)

```


## Control plots

### Maps

Three maps are generated to control successful merging of data sets.

```{r joined_climatology_control_PO4, fig.asp=0.6}
map_climatology(predictors, "PO4")
```

```{r joined_climatology_control_tem, fig.asp=0.6}
map_climatology(predictors, "tem")
```


### Predictor profiles

Likewise, predictor profiles for the North Atlantic (`r parameters$lat_Atl_profile` / `r parameters$lon_Atl_section`) are plotted to control successful merging of the data sets.

```{r predictor_profiles_N_Atl, fig.asp=1.5}

N_Atl <- predictors %>% 
  filter(lat == parameters$lat_Atl_profile,
         lon == parameters$lon_Atl_section)

N_Atl <- N_Atl %>% 
  select(-basin) %>% 
  pivot_longer(oxygen:gamma, names_to = "parameter", values_to = "value")

N_Atl %>% 
  ggplot(aes(value, depth)) +
  geom_path() +
  geom_point() +
  scale_y_reverse() +
  facet_wrap(~parameter,
             scales = "free_x",
             ncol = 2)

rm(N_Atl)

```

# Prepare predictor fields

## PO~4~*

### Calculation 

Currently, the predictor PO~4~* is calculated according to Clement and Gruber (2018), ie based on oxygen, as well as according to Gruber et al (2019), based on nitrate. Selection of one approach is done before mapping.

```{r calculate_phosphate_star_clement}

predictors <- predictors %>% 
  rename(phosphate = PO4) %>% 
  mutate(phosphate_star_oxy = phosphate + (oxygen / 170)  - 1.95,
         phosphate_star_nit = phosphate - NO3 / 16  + 2.9)

```

### Maps

```{r PO4_star_nit_climatology_maps, fig.asp=0.6}

map_climatology(predictors, "phosphate_star_nit")

```

```{r PO4_star_oxy_climatology_maps, fig.asp=0.6}

map_climatology(predictors, "phosphate_star_oxy")

```

### Sections

```{r PO4_star_nit_climatology_section, fig.asp=0.6}

section_climatology(predictors, "phosphate_star_nit")

```

```{r PO4_star_oxy_climatology_section, fig.asp=0.6}

section_climatology(predictors, "phosphate_star_oxy")

```


## AOU

### Calculation

AOU was calculated as the difference between saturation concentration and observed concentration.  

**CAVEAT**: Algorithms used to calculate oxygen saturation concentration are not yet identical in GLODAP data set (fitting) and predictor climatologies (mapping).

```{r calculate_aou_climatology}

predictors <- predictors %>% 
  mutate(oxygen_sat = gas_satconc(S = sal,
                                  t = tem,
                                  P = 1.013253,
                                  species = "O2"),
         aou = oxygen_sat - oxygen) %>% 
  select(-oxygen_sat)


```

### Maps

```{r aou_climatology_maps, fig.asp=0.6}

map_climatology(predictors, "aou")

```

### Sections

```{r aou_climatology_section, fig.asp=0.6}

section_climatology(predictors, "aou")

```

## Isoneutral slabs

The following boundaries for isoneutral slabs were defined:

- Atlantic: `r parameters$slabs_Atl`
- Indo-Pacific: `r parameters$slabs_Ind_Pac`

Continuous neutral density (gamma) values based on WOA18 are grouped into isoneutral slabs.

```{r cut_isoneutral_slabs}

predictors_Atl <- predictors %>% 
  filter(basin == "Atlantic") %>% 
  mutate(gamma_slab = cut(gamma, parameters$slabs_Atl))

predictors_Ind_Pac <- predictors %>% 
  filter(basin == "Indo-Pacific") %>% 
  mutate(gamma_slab = cut(gamma, parameters$slabs_Ind_Pac))

predictors <- bind_rows(predictors_Atl, predictors_Ind_Pac)

rm(predictors_Atl, predictors_Ind_Pac)

```

# Plot al predictor sections

Predictor sections along with lines are shown below for each (potential) predictor variable.

```{r predictors_observations_sections_map, fig.asp=0.6}

predictors %>%
  ggplot(aes(lon, lat)) +
  geom_bin2d(binwidth = c(1,1)) +
  geom_vline(xintercept = parameters$longitude_sections_regular,
             col = "white") +
  scale_fill_viridis_c(direction = -1) +
  coord_quickmap(expand = 0) +
  theme(legend.position = "bottom")

```

```{r all_predictor_sections, fig.asp=1}

vars <-
  c(
    "gamma",
    "sal",
    "tem",
    "phosphate",
    "phosphate_star_oxy",
    "phosphate_star_nit",
    "oxygen",
    "aou",
    "silicate",
    "NO3"
  )

for (i_var in vars) {
  print(section_climatology_regular(predictors, i_var))
}

```

# Write csv

```{r write_joined_predictor_fields}

predictors %>%
    write_csv(here::here("data/mapping/predictor_fields",
                         "W18_st_G16_opsn.csv"))

```


## WOA13 + GLODAP

Currently not used.

```{r join_WOA13_GLODAP_predictors, eval=FALSE}

predictors <- full_join(GLODAP_predictors, WOA13_gamma)
GLODAP_depths <- unique(GLODAP_predictors$depth)

rm(GLODAP_predictors, WOA13_gamma)

predictors <- predictors %>% 
  group_by(lat, lon) %>% 
  mutate(n_oxygen = sum(!is.na(oxygen)),
         n_PO4 = sum(!is.na(PO4)),
         n_silicate = sum(!is.na(silicate)),
         n_sal = sum(!is.na(sal)),
         n_tem = sum(!is.na(tem)),
         n_gamma = sum(!is.na(gamma))) %>% 
  ungroup()

predictors <- predictors %>% 
  filter(n_oxygen > 0,
         n_PO4 > 0,
         n_silicate > 0, 
         n_sal > 0, 
         n_tem > 0, 
         n_gamma > 0) %>% 
  select(-c(n_oxygen, 
            n_PO4, 
            n_silicate, 
            n_sal, 
            n_tem,
            n_gamma))

predictors <- predictors %>% 
  drop_na()

```



# Open tasks

- Check PO4* calculation
- Harmonize AOU calculation in fitting and mapping
- 

# Open questions

- 
