---
title: "Mapping Cstar"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(metR)
```

```{r read_parameters, include = FALSE}

parameters <-
  read_rds(here::here("data",
                       "parameters.rds"))
```

```{r read_mask_files}

basinmask <-
  read_csv(
    here::here(
      "data/World_Ocean_Atlas_2018/_summarized_files",
      "basin_mask_WOA18.csv"
    )
  )

basinmask_AIP <-
  read_csv(
    here::here(
      "data/World_Ocean_Atlas_2018/_summarized_files",
      "basin_mask_WOA18_AIP.csv"
    )
  )

landmask <-
  read_csv(
    here::here(
      "data/World_Ocean_Atlas_2018/_summarized_files",
      "land_mask_WOA18.csv"
    )
  )

```

```{r read_parameters_functions, include = FALSE}
source(here::here("code", "plotting_functions.R"))
```

```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

# Predictor fields

Currently, we use combined predictor fields:

- WOA18: S, T, and derived variables
- GLODAP16: Oxygen, PO4, NO3, Silicate, and derived variables

```{r read_predictor_file}

predictors <- 
    read_csv(here::here("data/mapping/predictor_fields",
                         "W18_st_G16_opsn.csv"))

```


# Load MLR models

```{r load_eMLR_models}

lm_all_wide_Cstar <-
  read_csv(here::here("data/eMLR",
                       "lm_all_wide_Cstar.csv"))

```

# Merge MLRs + climatologies

```{r merge_model_coeff_predictor_climatology}

lm_all_wide_Cstar <- lm_all_wide_Cstar %>% 
  mutate(model = str_remove(model, "Cstar ~ "))
         
Cstar <- full_join(predictors, lm_all_wide_Cstar)

rm(predictors, lm_all_wide_Cstar)

```


# Map Cstar

## Apply MLRs to predictor

```{r calculate_Cstar, eval=FALSE}

Cstar <- Cstar %>% 
  mutate(Cstar =
          `coeff_(Intercept)` +
           coeff_aou * aou +
           coeff_oxygen * oxygen +
           coeff_phosphate * phosphate +
           coeff_phosphate_star * phosphate_star +
           coeff_silicate * silicate +
           coeff_sal * sal + 
           coeff_tem * tem)

Cstar <- Cstar %>% 
  select(lon, lat, depth, era, basin, Cstar, gamma)

```

```{r Cstar_average, eval=FALSE}

Cstar_average <- Cstar %>%
  group_by(lon, lat, depth, era, basin) %>%
  summarise(Cstar_mean = mean(Cstar, na.rm = TRUE),
            Cstar_sd = sd(Cstar, na.rm = TRUE),
            gamma_mean = mean(gamma, na.rm = TRUE)) %>%
  ungroup()

rm(Cstar)

Cstar_average_Atl <- Cstar_average %>% 
  filter(basin == "Atlantic") %>% 
  mutate(gamma_slab = cut(gamma_mean, parameters$slabs_Atl))

Cstar_average_Ind_Pac <- Cstar_average %>% 
  filter(basin == "Indo-Pacific") %>% 
  mutate(gamma_slab = cut(gamma_mean, parameters$slabs_Ind_Pac))

Cstar_average <- bind_rows(Cstar_average_Atl, Cstar_average_Ind_Pac)

rm(Cstar_average_Atl, Cstar_average_Ind_Pac)

```

## Mean Cstar sections

For each basin and era combination, the zonal mean Cstar is calculated. Likewise, sd is calculated for the averaging of the mean basin fields.

```{r Calculate_Cstar_mean_sections, eval=FALSE}

Cstar_average <- left_join(Cstar_average,
                           basinmask_AIP %>% select(-basin))

Cstar_average_zonal <- Cstar_average %>%
  group_by(lat, depth, era, basin, basin_AIP) %>%
  summarise(across(
    c(
      "Cstar_mean",
      "Cstar_sd",
      "gamma_mean"
    ),
    list(mean = ~ mean(.x, na.rm = TRUE),
         sd = ~ sd(.x, na.rm = TRUE))
  )) %>%
  ungroup()


Cstar_average_zonal_Atl <- Cstar_average_zonal %>% 
  filter(basin == "Atlantic") %>% 
  mutate(gamma_slab = cut(gamma_mean_mean, parameters$slabs_Atl))

Cstar_average_zonal_Ind_Pac <- Cstar_average_zonal %>% 
  filter(basin == "Indo-Pacific") %>% 
  mutate(gamma_slab = cut(gamma_mean_mean, parameters$slabs_Ind_Pac))

Cstar_average_zonal <- bind_rows(Cstar_average_zonal_Atl, Cstar_average_zonal_Ind_Pac)

rm(Cstar_average_zonal_Atl, Cstar_average_zonal_Ind_Pac)

```

# Write csv

```{r write_Cstar_files, eval=FALSE}

Cstar_average %>%
    write_csv(here::here("data/mapping/_summarized_files",
                         "Cstar_average.csv"))

Cstar_average_zonal %>%
    write_csv(here::here("data/mapping/_summarized_files",
                         "Cstar_average_zonal.csv"))

```
