---
title: "World Ocean Atlas 2013 - D. Clement"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidync)
library(reticulate)
library(oce)
```

```{r read_parameters, include = FALSE}
parameters <-
  read_rds(here::here("data",
                       "parameters.rds"))
```

```{r read_mask_files}

basinmask <- read_csv(here::here("data/World_Ocean_Atlas_2018/_summarized_files",
                                 "basin_mask_WOA18.csv"))

landmask <- read_csv(here::here("data/World_Ocean_Atlas_2018/_summarized_files",
                                 "land_mask_WOA18.csv"))

```

```{r read_functions, include = FALSE}
source(here::here("code", "plotting_functions.R"))
```

```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

# Data source

Dominic Clement provided a netcdf file with the basin mask and neutral densities used in Clement and Gruber (2018), both derived from the World Ocean Atlas 2013.

# Read ncdfs

```{r read_WOA13_dclement}


nd_mask <- tidync(here::here("data/World_Ocean_Atlas_2013_Clement",
                             "nd_mask.nc"))

nd_mask_tibble <- nd_mask %>% hyper_tibble()

nd_mask_tibble <- nd_mask_tibble %>%
  mutate(gamma = if_else(gamma == -999, NaN, gamma)) %>%
  drop_na()

nd_mask_tibble <- nd_mask_tibble %>%
  rename(lat = latitude,
         lon = longitude) %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon))



```

# Apply basin mask


# Plots

Below, following subsets of the climatologies are plotted for all relevant parameters:  

- Horizontal planes at `r parameters$depth_levels`m
- Meridional sections at longitudes: `r parameters$longitude_sections_basin`

Section locations are indicated as white lines in maps.

## Basin map

Follwoing basin mask is not further used in this project.

```{r basin_map_WOA13_dclement, fig.asp=0.6}

nd_mask_tibble %>%
  filter(depth == 0) %>%
  ggplot(aes(lon, lat, fill = as.factor(mask))) +
  geom_raster() +
  geom_vline(
    xintercept = c(
      parameters$lon_Atl_section,
      parameters$lon_Pac_section,
      parameters$lon_Ind_section
    ),
    col = "white"
  ) +
  coord_quickmap(expand = 0) +
  scale_fill_brewer(palette = "Set1",
                    name = "basin mask") +
  theme(legend.position = "top",
        axis.title = element_blank())

```

## Neutral density

### Maps

```{r neutral_density_surface_WOA18_map}

map_climatology(nd_mask_tibble, "gamma")

```

### Sections

```{r neutral_density_sections_WOA18, fig.asp=0.8}

section_climatology(nd_mask_tibble, "gamma")

```

### Sections shallow

```{r neutral_density_sections_WOA18_shallow, fig.asp=0.8}

section_climatology_shallow(nd_mask_tibble, "gamma")

```


## Write file

```{r write_WOA13_dclement_file}

nd_mask_tibble %>%
  write_csv(here::here("data/World_Ocean_Atlas_2013_Clement/_summarized_files",
                       "WOA13_mask_gamma.csv"))

rm(nd_mask, nd_mask_tibble, Atl_lon, Pac_lon, depth_surface_selection)

```
