---
title: "GLODAPv2_2020"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(patchwork)
```


```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

# Read master file

Main data source for this project is `GLODAPv2.2020_Merged_Master_File.csv` downloaded from [glodap.info](https://www.glodap.info/) in June 2020.

```{r read_GLODAPv2_2020_merged_master_file}

GLODAP <- read_csv(here::here("data/GLODAPv2_2020/Merged_data_product",
                              "GLODAPv2.2020_Merged_Master_File.csv"),
                   na = "-9999",
                   col_types = cols(.default = col_double()))

# relevant columns
GLODAP <- GLODAP %>% 
  select(cruise:talkqc)

GLODAP <- GLODAP %>% 
  mutate(date = ymd(paste(year, month, day))) %>% 
         #decade = as.factor(floor(year / 10) * 10)) %>% 
  relocate(date)

GLODAP <- GLODAP %>% 
  select(-c(month:minute, 
            maxsampdepth, pressure,
            theta, sigma0:sigma4,
            nitrite:nitritef))

```

# Data preparation

## Flags and missing data

```{r set_criteria_flagging}

flag_f <- 2
flag_qc <- 1

```

Only rows (samples) for which all relevant parameters are available were selected.
In additions, following flagging criteria were applied:

- flag f:     `r flag_f`
- flag qc:    `r flag_qc`
  
Summary statistics were calculated during cleaning process.

```{r clean_GLODAP_master_file}

GLODAP_stats <- GLODAP %>% 
  summarise(total = n())

##

GLODAP <- GLODAP %>% 
  filter(!is.na(tco2))

GLODAP <- GLODAP %>% 
  filter(tco2f == flag_f) %>% 
  select(-tco2f)

GLODAP <- GLODAP %>% 
  filter(tco2qc == flag_qc) %>% 
  select(-tco2qc)

GLODAP_stats_temp <- GLODAP %>% 
  summarise(tco2 = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP <- GLODAP %>% 
  filter(!is.na(talk))

GLODAP <- GLODAP %>% 
  filter(talkf == flag_f) %>% 
  select(-talkf)

GLODAP <- GLODAP %>% 
  filter(talkqc == flag_qc) %>% 
  select(-talkqc)

##

GLODAP <- GLODAP %>% 
  filter(!is.na(phosphate))

GLODAP <- GLODAP %>% 
  filter(phosphatef == flag_f) %>% 
  select(-phosphatef)

GLODAP <- GLODAP %>% 
  filter(phosphateqc == flag_qc) %>% 
  select(-phosphateqc)

GLODAP_stats_temp <- GLODAP %>% 
  summarise(C_star_variables = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP <- GLODAP %>% 
  filter(!is.na(temperature))

##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(salinity))

GLODAP <- GLODAP %>% 
  filter(salinityf == flag_f) %>% 
  select(-salinityf)

GLODAP <- GLODAP %>% 
  filter(salinityqc == flag_qc) %>% 
  select(-salinityqc)

##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(silicate))

GLODAP <- GLODAP %>% 
  filter(silicatef == flag_f) %>% 
  select(-silicatef)

GLODAP <- GLODAP %>% 
  filter(silicateqc == flag_qc) %>% 
  select(-silicateqc)

##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(oxygen))

GLODAP <- GLODAP %>% 
  filter(oxygenf == flag_f) %>% 
  select(-oxygenf)

GLODAP <- GLODAP %>% 
  filter(oxygenqc == flag_qc) %>% 
  select(-oxygenqc)

##

GLODAP <- GLODAP %>% 
  filter(!is.na(aou))

GLODAP <- GLODAP %>% 
  filter(aouf == flag_f) %>% 
  select(-aouf)

##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(nitrate))

GLODAP <- GLODAP %>% 
  filter(nitratef == flag_f) %>% 
  select(-nitratef)

GLODAP <- GLODAP %>% 
  filter(nitrateqc == flag_qc) %>% 
  select(-nitrateqc)

##

GLODAP <- GLODAP %>% 
  filter(!is.na(depth))

GLODAP <- GLODAP %>% 
  filter(!is.na(gamma))

##

GLODAP_stats_temp <- GLODAP %>% 
  summarise(eMLR_variables = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)

rm(GLODAP_stats_temp)
rm(flag_f, flag_qc)

```

## Spatial boundaries

```{r set_criteria_spatial_boundaries}

min_depth <- 150
min_bottomdepth <- 500
max_lat <- 65

```

Only observations with:

- minimum sampling depth:  `r min_depth`m
- minimum bottom depth:  `r min_bottomdepth`m
- maximum latitude:  `r max_lat`°N

were taken into consideration.

```{r apply_spatial_boundaries}

GLODAP <- GLODAP %>% 
  filter(depth >= min_depth)

GLODAP_stats_temp <- GLODAP %>% 
  summarise(eMLR_variables_150m = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP <- GLODAP %>% 
  filter(latitude <= max_lat)

GLODAP <- GLODAP %>% 
  filter(bottomdepth >= min_bottomdepth)

GLODAP_stats_temp <- GLODAP %>% 
  summarise(eMLR_variables_150m_65N_500m = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)


GLODAP_stats_long <- GLODAP_stats %>% 
  pivot_longer(1:length(GLODAP_stats), names_to = "parameter", values_to = "n")

GLODAP_stats_long  %>%  write_csv(here::here("data/GLODAPv2_2020/_summarized_data_files",
                                             "GLODAPv2.2020_stats.csv"))

rm(GLODAP_stats_long, GLODAP_stats)
rm(min_depth, max_lat, min_bottomdepth)

```

## Reference eras

```{r set_criteria_eras}

start <- 1981
JGOFS <- 1999
GOSHIP <- 2012

```

Samples were assigned to following eras:

- JGOFS-WOCE: `r start` - `r JGOFS`
- GO-SHIP:    `r JGOFS+1` - `r GOSHIP`
- new-era:    > `r GOSHIP+1`


```{r assign_eras}

GLODAP <- GLODAP %>%
  filter(year >= start) %>% 
  mutate(era = "JGOFS_WOCE",
         era = if_else(year > JGOFS, "GO_SHIP", era),
         era = if_else(year > GOSHIP, "new_era", era))

```

## Horizontal gridding

For merging with other data sets, all observations were grouped into latitude intervals of:

- 1° x 1°

```{r grid_spatially_1x1}

GLODAP <- GLODAP %>% 
  mutate(lat = cut(latitude, seq(-90, 90, 1), seq(-89.5, 89.5, 1)),
         lat = as.numeric(as.character(lat)),
         lon = cut(longitude, seq(-180, 180, 1), seq(-179.5, 179.5, 1)),
         lon = as.numeric(as.character(lon))) %>% 
  arrange(date) %>% 
  select(-c(latitude,longitude))

```


## Basin mask

The basin mask from the World Ocean Atlas was used. For details consult the data base section for WOA18 data [Link](/read_World_Ocean_Atlas_2018.Rmd).

```{r read_basin_mask_file}

basinmask <- read_csv(here::here("data/World_Ocean_Atlas_2018/_summarized_files",
                                 "basin_mask_WOA18.csv"))

```

```{r WOA_basin_mask_GLODAP_observations_map}

GLODAP_obs <- GLODAP %>% 
  group_by(lat, lon) %>% 
  summarise(n = n()) %>% 
  ungroup()

mapWorld <- borders("world", col = "gray60", fill = "gray60") 

ggplot() +
  #mapWorld +
  geom_raster(data = basinmask, aes(lon, lat, fill = basin)) +
  geom_raster(data = GLODAP_obs, aes(lon, lat)) +
  scale_fill_brewer(palette = "Dark2") +
  coord_quickmap(expand = 0)

rm(mapWorld, GLODAP_obs)

```

Please note that some GLODAP observations were made outside the WOA18 basin mask and will be removed for further analysis.

```{r join_basin_mask}

GLODAP <- inner_join(GLODAP, basinmask)
rm(basinmask)

```


## Write summary file 
 
```{r write_clean_data_file}

GLODAP  %>%  write_csv(here::here("data/GLODAPv2_2020/_summarized_data_files",
                                  "GLODAPv2.2020_clean.csv"))

rm(GLODAP)

```


# Overview plots

## Cleaning stats

```{r GLODAP_cleaning_stats}

GLODAP_stats_long <-  read_csv(here::here("data/GLODAPv2_2020/_summarized_data_files",
                                             "GLODAPv2.2020_stats.csv"))

GLODAP_stats_long <- GLODAP_stats_long %>%
  mutate(parameter = fct_reorder(parameter, n))

GLODAP_stats_long %>% 
  ggplot(aes(parameter, n/1000)) +
  geom_col() +
  coord_flip() +
  labs(y = "n (1000)") +
  theme(axis.title.y = element_blank())

rm(GLODAP_stats_long)

```


## Assign coarse spatial grid

For the following plots, the cleaned data set was re-opened and observations were gridded spatially to 5°x5° intervals.


```{r read_GLODAP_clean}

GLODAP <- read_csv(here::here("data/GLODAPv2_2020/_summarized_data_files",
                              "GLODAPv2.2020_clean.csv"))

```

```{r grid_spatially_5x5}

GLODAP <- GLODAP %>% 
  mutate(lat_grid = cut(lat, seq(-90, 90, 5), seq(-87.5, 87.5, 5)),
         lat_grid = as.numeric(as.character(lat_grid)),
         lon_grid = cut(lon, seq(-180, 180, 5), seq(-177.5, 177.5, 5)),
         lon_grid = as.numeric(as.character(lon_grid))) %>% 
  arrange(date)

```


## Histogram Zonal coverage

```{r coverage_histogram_zonal}

GLODAP_histogram_lat <- GLODAP %>% 
  group_by(era, lat_grid, basin) %>% 
  tally() %>% 
  ungroup()

GLODAP_histogram_lat %>% 
  ggplot(aes(lat_grid, n, fill = era)) +
  geom_col() +
  scale_x_continuous(breaks = seq(-87.5,90,10)) +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap(~basin) +
  coord_flip(expand = 0)

rm(GLODAP_histogram_lat)

```

## Histogram temporal coverage

```{r coverage_histogram_temporal}

GLODAP_histogram_year <- GLODAP %>% 
  group_by(year, basin) %>% 
  tally() %>% 
  ungroup()

era_median_year <- GLODAP %>% 
  group_by(era) %>% 
  summarise(t_ref = median(year)) %>% 
  ungroup()

GLODAP_histogram_year %>% 
  ggplot() +
  geom_vline(xintercept = c(JGOFS+0.5, GOSHIP+0.5)) +
  geom_col(aes(year, n, fill=basin)) +
  geom_point(data = era_median_year, aes(t_ref,0, shape="Median year"), size=2, fill="white") +
  scale_fill_brewer(palette = "Dark2") +
  scale_shape_manual(values = 24) +
  coord_cartesian(expand = 0)

rm(GLODAP_histogram_year,
   era_median_year)

```

## Zonal temporal coverage (Hovmoeller)

```{r coverage_hovmoeller, fig.asp=1}

GLODAP_hovmoeller_year <- GLODAP %>% 
  group_by(year, lat_grid, basin) %>% 
  tally() %>% 
  ungroup()

GLODAP_hovmoeller_year %>% 
  ggplot(aes(year, lat_grid, fill=log10(n))) +
  geom_tile() +
  geom_vline(xintercept = c(1999.5, 2012.5)) +
  scale_fill_viridis_c(option = "magma", direction = -1) +
  facet_wrap(~basin, ncol=1)

rm(GLODAP_hovmoeller_year)

```

## Coverage maps by era

```{r coverage_maps_era, fig.asp=1.5}

mapWorld <- borders("world", colour="gray60", fill="gray60")


GLODAP <- GLODAP %>% 
  mutate(era = factor(era, c("JGOFS_WOCE", "GO_SHIP", "new_era")))

GLODAP %>%
  ggplot(aes(lon, lat)) +
  mapWorld +
  geom_bin2d(binwidth = c(1,1)) +
  scale_fill_viridis_c(option = "magma", direction = -1, trans = "log10") +
  scale_x_continuous(breaks = seq(-180, 180, 30)) +
  scale_y_continuous(breaks = seq(-90, 90, 30)) +
  coord_quickmap(expand = FALSE) +
  facet_wrap(~era, ncol = 1)

rm(mapWorld)

```

# Individual cruise sections

Zonal and meridional section plots are produce for each cruise individually and stored locally.

```{r individual_cruises_clean_triplicates, eval = FALSE}

cruises <- GLODAP %>% 
  group_by(cruise) %>% 
  summarise(date_mean = mean(date, na.rm = TRUE),
            n = n()) %>% 
  ungroup() %>% 
  arrange(date_mean)

GLODAP <- full_join(GLODAP, cruises)

mapWorld <- borders("world", colour="gray60", fill="gray60")

n <- 0
for (i_cruise in unique(cruises$cruise)) {

#i_cruise <- unique(cruises$cruise)[1]
n <- n+1
print(n)  
  
GLODAP_cruise <- GLODAP %>%
  filter(cruise == i_cruise) %>% 
  arrange(date)

cruises_cruise <- cruises %>%
  filter(cruise == i_cruise)
  
map <- GLODAP_cruise %>%
  ggplot(aes(lon, lat)) +
  mapWorld+
  geom_point(aes(col=date)) +
  geom_path() +
  coord_quickmap(expand = FALSE) +
  scale_color_viridis_c(trans = "date") +
  labs(title = paste("Mean date:", cruises_cruise$date_mean,
                     "| cruise:", cruises_cruise$cruise,
                     "| n(samples):", cruises_cruise$n))


lon_section <- GLODAP_cruise %>%
  ggplot(aes(lon, depth)) +
  scale_y_reverse() +
  scale_color_viridis_c()

lon_tco2 <- lon_section+
  geom_point(aes(col=tco2))

lon_talk <- lon_section+
  geom_point(aes(col=talk))

lon_phosphate <- lon_section+
  geom_point(aes(col=phosphate))


lat_section <- GLODAP_cruise %>%
  ggplot(aes(lat, depth)) +
  scale_y_reverse() +
  scale_color_viridis_c()

lat_tco2 <- lat_section+
  geom_point(aes(col=tco2))

lat_talk <- lat_section+
  geom_point(aes(col=talk))

lat_phosphate <- lat_section+
  geom_point(aes(col=phosphate))

map /
((lat_tco2 / lat_talk / lat_phosphate) |
(lon_tco2 / lon_talk / lon_phosphate))

ggsave(here::here("output/figure/data/all_cruises_clean",
                  paste("GLODAP_cruise_date",
                        cruises_cruise$date_mean,
                        "n",
                        cruises_cruise$n,
                        "cruise",
                        cruises_cruise$cruise,
                        ".png",
                        sep = "_")),
                  width = 9, height = 9)

rm(map,
   lon_section, lat_section,
   lat_tco2, lat_talk, lat_phosphate, lon_tco2, lon_talk, lon_phosphate,
   GLODAP_cruise, cruises_cruise)

}


# library("rnaturalearth")
# library("rnaturalearthdata")
# library("sf")
# 
# world <- ne_countries(scale = "small", returnclass = "sf")
# class(world)
# 
# GLODAP_map <- GLODAP %>% 
#   group_by(lat_grid, lon_grid) %>% 
#   tally() %>% 
#   ungroup()
# 
# ggplot() +
#   geom_raster(data = GLODAP_map, aes(lon_grid, lat_grid)) +
#   geom_sf(data = world) +
#   coord_sf(crs = "+proj=robin +lat_0=0 +lon_0=0 +x0=0 +y0=0")


# https://gist.github.com/clauswilke/783e1a8ee3233775c9c3b8bfe531e28a
# https://twitter.com/clauswilke/status/1066024436208406529
# https://www.r-spatial.org/r/2018/10/25/ggplot2-sf.html
```

# Open tasks

- move A16 cruises from new_era to GO-SHIP era

# Questions
