---
title: "GLODAPv2_2020"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(patchwork)
library(cutr)
```

```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

```{r read_parameters, include = FALSE}
parameters <-
  read_rds(here::here("data",
                       "parameters.rds"))
```

```{r read_mask_files, include = FALSE}

basinmask <- read_csv(
  here::here(
    "data/World_Ocean_Atlas_2018/_summarized_files",
    "basin_mask_WOA18.csv"
  )
)

landmask <-  read_csv(
  here::here(
    "data/World_Ocean_Atlas_2018/_summarized_files",
    "land_mask_WOA18.csv"
  )
)
```


# Read files

Main data source for this project is `GLODAPv2.2020_Merged_Master_File.csv` downloaded from [glodap.info](https://www.glodap.info/) in June 2020.

```{r read_GLODAPv2_2020_merged_master_file}

GLODAP <-
  read_csv(
    here::here(
      "data/GLODAPv2_2020/Merged_data_product",
      "GLODAPv2.2020_Merged_Master_File.csv"
    ),
    na = "-9999",
    col_types = cols(.default = col_double())
  )


# select relevant columns
GLODAP <- GLODAP %>%
  select(cruise:talkqc)

# create date column
GLODAP <- GLODAP %>%
  mutate(date = ymd(paste(year, month, day))) %>%
  #decade = as.factor(floor(year / 10) * 10)) %>%
  relocate(date)

# harmonize column names
GLODAP <- GLODAP  %>%
  rename(sal = salinity,
         tem = temperature)

# harmonize coordinates
GLODAP <- GLODAP  %>%
  rename(lon = longitude,
         lat = latitude) %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon))

# remove irrelevant columns
GLODAP <- GLODAP %>%
  select(-c(month:minute,
            maxsampdepth, bottle, sigma0:sigma4,
            nitrite:nitritef))

GLODAP <- GLODAP %>%
  mutate_at(vars(ends_with(c("f", "qc"))), as.factor)

```

# Exploratory data analysis

```{r EDA_GLODAP_raw, eval=FALSE}

source("code/eda.R")
eda(GLODAP, "GLODAPv2_2020")
rm(eda)

```

The output of an automated Exploratory Data Analysis (EDA) performed with the package `DataExplorer` can be accessed here:

[Link to EDA report of GLODAPv2_2020 raw data](https://jens-daniel-mueller.github.io/Cant_eMLR/EDA_report_GLODAPv2_2020.html){target="_blank}


# Data preparation

## Horizontal gridding

For merging with other data sets, all observations were grouped into latitude intervals of:

- 1° x 1°

```{r grid_spatially_1x1}

GLODAP <- GLODAP %>% 
  mutate(lat = cut(lat, seq(-90, 90, 1), seq(-89.5, 89.5, 1)),
         lat = as.numeric(as.character(lat)),
         lon = cut(lon, seq(20, 380, 1), seq(20.5, 379.5, 1)),
         lon = as.numeric(as.character(lon)))


```


## Reference eras

Samples were assigned to following eras:

- JGOFS_WOCE:
`r parameters$year_JGOFS_start` - `r parameters$year_JGOFS_end`

- GO_SHIP:
`r parameters$year_JGOFS_end+1` - `r parameters$year_GOSHIP_end`

- new_era:    > `r parameters$year_GOSHIP_end+1`


```{r assign_eras}

GLODAP <- GLODAP %>%
  filter(year >= parameters$year_JGOFS_start) %>% 
  mutate(era = "JGOFS_WOCE",
         era = if_else(year > parameters$year_JGOFS_end, "GO_SHIP", era),
         era = if_else(year > parameters$year_GOSHIP_end, "new_era", era))

GLODAP <- GLODAP %>% 
  mutate(era = factor(era, c("JGOFS_WOCE", "GO_SHIP", "new_era")))


```


## Spatial boundaries

### Depth

Only observations were taken into consideration with:

- minimum sampling depth:  `r parameters$depth_min`m

```{r apply_depth_min}

# create observations grid and summary stats before remving data

GLODAP_obs_grid <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "all")

GLODAP_stats <- GLODAP %>% 
  summarise(all = n())

##

GLODAP <- GLODAP %>% 
  filter(depth >= parameters$depth_min)

```

### Bottomdepth

Only observations were taken into consideration with:

- minimum bottom depth:  `r parameters$bottomdepth_min`m


```{r apply_bottomdepth_min}

GLODAP <- GLODAP %>% 
  filter(bottomdepth >= parameters$bottomdepth_min)


```


### Basin mask

The basin mask from the World Ocean Atlas was used. For details consult the data base subsection for [WOA18](https://jens-daniel-mueller.github.io/Cant_eMLR/read_World_Ocean_Atlas_2018.html) data.

Please note that some GLODAP observations were made outside the WOA18 basin mask and will be removed for further analysis (eg North Sea, Sea of Japan).

```{r join_GLODAP_basin_mask}

GLODAP <- inner_join(GLODAP, basinmask)

##

GLODAP_stats_temp <- GLODAP %>% 
  summarise(spatial_boundaries = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "spatial_boundaries")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)


```


```{r WOA18_basin_mask_GLODAP_observations_map}

GLODAP_obs <- GLODAP %>% 
  group_by(lat, lon) %>% 
  summarise(n = n()) %>% 
  ungroup()

ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat), fill = "grey80") +
  geom_raster(data = basinmask, aes(lon, lat, fill = basin)) +
  geom_raster(data = GLODAP_obs, aes(lon, lat)) +
  scale_fill_brewer(palette = "Dark2") +
  coord_quickmap(expand = 0) +
  theme(legend.position = "top",
        legend.title = element_blank(),
        axis.title = element_blank())

rm(GLODAP_obs)

```




## Flags and missing data

Only rows (samples) for which all relevant parameters are available were selected, ie NA's were removed.

According to Olsen et al (2020), flags within the merged master file identify:

- f:  

  - 2: Acceptable
  - 0: Interpolated or calculated value (or Average of replicate, Manual chromatographic peak measurement)
  - 9: Data not used (so, only NA data should have this flag)
  
  
- qc:  

  - 1: Adjusted or unadjusted data
  - 0: Data appear of good quality but have not been subjected to full secondary QC
  - data with poor or uncertain quality are excluded.

Following flagging criteria were taken into account:

- flag_f:     `r parameters$flag_f`
- flag_qc:    `r parameters$flag_qc`

Summary statistics were calculated during cleaning process.

### tco2

#### NA

The vast majority of rows is removed due to missing tco2 observations.

```{r tco2_na_subset}

GLODAP <- GLODAP %>% 
  filter(!is.na(tco2))

##

GLODAP_stats_temp <- GLODAP %>% 
  summarise(tco2_values = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "tco2_values")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)

```

#### f flag

```{r tco2_f_subset, fig.asp=1}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, tco2f)

ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat),
              fill = "grey80") +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ tco2f) +
  coord_quickmap(expand = 0) +
  theme(legend.position = "top")

###

GLODAP <- GLODAP %>%
  filter(tco2f %in% parameters$flag_f)

# ##
# 
# GLODAP_stats_temp <- GLODAP %>%
#   summarise(tco2_f_flag = n())
# 
# GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
# rm(GLODAP_stats_temp)
# 
# ##
# 
# GLODAP_obs_grid_temp <- GLODAP %>%
#   count(lat, lon, era) %>%
#   mutate(cleaning_level = "tco2_f_flag")
# 
# GLODAP_obs_grid <-
#   bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)
# 
# rm(GLODAP_obs_grid_temp)

```


#### qc flag

```{r tco2_qc_subset, fig.asp=1}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, tco2qc)

ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat),
              fill = "grey80") +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ tco2qc) +
  coord_quickmap(expand = 0) +
  theme(legend.position = "top")

###

GLODAP <- GLODAP %>%
  filter(tco2qc %in% parameters$flag_qc)

##

GLODAP_stats_temp <- GLODAP %>%
  summarise(tco2_qc_flag = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era) %>%
  mutate(cleaning_level = "tco2_qc_flag")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)

```



### talk

#### NA

Quite a few tco2 observations are discarded, due to missing talk data, in particular in the JGOFS/WOCE era. However, there seems to be a high number of observations remaining from the affected cruises, so that the coverage does not seem to be reduced drastically.

```{r talk_na_subset, fig.asp=1}

GLODAP <- GLODAP %>% 
  mutate(talkna = if_else(is.na(talk), "NA", "Value"))

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, talkna)

ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat),
              fill = "grey80") +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ talkna) +
  coord_quickmap(expand = 0) +
  theme(legend.position = "top")

GLODAP <- GLODAP %>% 
  select(-talkna) %>% 
  filter(!is.na(talk))

##

GLODAP_stats_temp <- GLODAP %>% 
  summarise(talk_values = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "talk_values")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)


```

#### f flag

Restricting the f flag to 2 would results in data gaps in the south-east Pacific. Interpolated or calculated data (f flag 0) are therefore included.

```{r talk_f_subset, fig.asp=1}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, talkf)

ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat), fill = "grey80") +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
    scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era~talkf) +
  coord_quickmap(expand = 0) +
  theme(legend.position = "top",
        legend.title = element_blank())


# ###

GLODAP <- GLODAP %>%
  filter(talkf %in% parameters$flag_f)

##
# 
# GLODAP_stats_temp <- GLODAP %>%
#   summarise(talk_f_flag = n())
# 
# GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
# rm(GLODAP_stats_temp)
# 
# ##
# 
# GLODAP_obs_grid_temp <- GLODAP %>%
#   count(lat, lon, era) %>%
#   mutate(cleaning_level = "talk_f_flag")
# 
# GLODAP_obs_grid <-
#   bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)
# 
# rm(GLODAP_obs_grid_temp)

```

#### qc flag

```{r talk_qc_subset, fig.asp=1}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, talkqc)

ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat), fill = "grey80") +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
    scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era~talkqc) +
  coord_quickmap(expand = 0) +
  theme(legend.position = "top",
        legend.title = element_blank())


###

GLODAP <- GLODAP %>%
  filter(talkqc %in% parameters$flag_qc)

##

GLODAP_stats_temp <- GLODAP %>%
  summarise(talk_qc_flag = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era) %>%
  mutate(cleaning_level = "talk_qc_flag")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)

```


### Phosphate

#### NA

Quite a few tco2/talk observations are discarded, due to missing phosphate data. However, there seems to be a high number of observations remaining from the affected cruises, so that the coverage does not seem to be reduced drastically.

```{r phosphate_na_subset, fig.asp=1}

GLODAP <- GLODAP %>% 
  mutate(phosphatena = if_else(is.na(phosphate), "NA", "Value"))

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, phosphatena)

ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat),
              fill = "grey80") +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ phosphatena) +
  coord_quickmap(expand = 0) +
  theme(legend.position = "top")

GLODAP <- GLODAP %>% 
  select(-phosphatena) %>% 
  filter(!is.na(phosphate))

##

GLODAP_stats_temp <- GLODAP %>% 
  summarise(phosphate_values = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "phosphate_values")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)


```

#### f flag

Restricting the f flag to 2 would not drastically reduce available data. Interpolated or calculated data (f flag 0) are therefore included.

```{r phosphate_f_subset, fig.asp=1}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, phosphatef)

ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat), fill = "grey80") +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
    scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era~phosphatef) +
  coord_quickmap(expand = 0) +
  theme(legend.position = "top",
        legend.title = element_blank())


###

GLODAP <- GLODAP %>%
  filter(phosphatef %in% parameters$flag_f)

# ##
# 
# GLODAP_stats_temp <- GLODAP %>%
#   summarise(phosphate_f_flag = n())
# 
# GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
# rm(GLODAP_stats_temp)
# 
# ##
# 
# GLODAP_obs_grid_temp <- GLODAP %>%
#   count(lat, lon, era) %>%
#   mutate(cleaning_level = "phosphate_f_flag")
# 
# GLODAP_obs_grid <-
#   bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)
# 
# rm(GLODAP_obs_grid_temp)

```

#### qc flag

Phosphate data for which secondary quality was not applied (qc flag 0) were removed.

```{r phosphate_qc_subset, fig.asp=1}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, phosphateqc)

ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat), fill = "grey80") +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
    scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era~phosphateqc) +
  coord_quickmap(expand = 0) +
  theme(legend.position = "top",
        legend.title = element_blank())


###

GLODAP <- GLODAP %>%
  filter(phosphateqc %in% parameters$flag_qc)

##

GLODAP_stats_temp <- GLODAP %>%
  summarise(phosphate_qc_flag = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era) %>%
  mutate(cleaning_level = "phosphate_qc_flag")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)

```



### eMLR variables

Rows with missing eMLR variables were removed, only the qc flag was considered.

```{r eMLR_subset}

GLODAP <- GLODAP %>% 
  filter(!is.na(tem))

##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(sal))

GLODAP <- GLODAP %>%
  filter(salinityf %in% parameters$flag_f)

GLODAP <- GLODAP %>%
  filter(salinityqc %in% parameters$flag_qc)

##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(silicate))

GLODAP <- GLODAP %>%
  filter(silicatef %in% parameters$flag_f)

GLODAP <- GLODAP %>%
  filter(silicateqc %in% parameters$flag_qc)
##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(oxygen))

GLODAP <- GLODAP %>%
  filter(oxygenf %in% parameters$flag_f)

GLODAP <- GLODAP %>%
  filter(oxygenqc %in% parameters$flag_qc)
##

GLODAP <- GLODAP %>% 
  filter(!is.na(aou))

GLODAP <- GLODAP %>%
  filter(aouf %in% parameters$flag_f)
##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(nitrate))

GLODAP <- GLODAP %>%
  filter(nitratef %in% parameters$flag_f)

GLODAP <- GLODAP %>%
  filter(nitrateqc %in% parameters$flag_qc)
##

GLODAP <- GLODAP %>% 
  filter(!is.na(depth))

GLODAP <- GLODAP %>% 
  filter(!is.na(gamma))

##

GLODAP_stats_temp <- GLODAP %>% 
  summarise(eMLR_variables = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)

rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "eMLR_variables")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)

```


### Manual adjustment A16 cruise

For harmonization with Gruber et al. (2019), cruises 1041 (A16N) and 1042 (A16S) were grouped into the GO_SHIP area despite taking place in 2013/14.

```{r GLODAP_Atl_cruises_2013_2014_map}

GLODAP_cruises <- GLODAP %>% 
  filter(basin == "Atlantic",
         year %in% c(2013, 2014)) %>% 
  count(lat, lon, cruise)

ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat), fill = "grey80") +
  geom_raster(data = GLODAP_cruises, aes(lon, lat, fill = as.factor(cruise))) +
  scale_fill_brewer(palette = "Dark2") +
  coord_quickmap(expand = 0) +
  theme(legend.position = "top",
        legend.title = element_blank())

rm(GLODAP_cruises)

```


```{r switch_eras_manually}

GLODAP <- GLODAP %>%
  mutate(era = as.character(era)) %>% 
  mutate(era = if_else(cruise %in% c(1041, 1042),
                      "GO_SHIP", era))

GLODAP <- GLODAP %>% 
  mutate(era = factor(era, c("JGOFS_WOCE", "GO_SHIP", "new_era")))

```

## Create clean observations grid

```{r create_clean_obs_grid}

GLODAP_obs_grid_clean <- GLODAP %>% 
  count(lat, lon) %>% 
  select(-n)

```



## Write summary file
 
```{r write_clean_data_files}

GLODAP_stats_long <- GLODAP_stats %>%
  pivot_longer(1:length(GLODAP_stats),
               names_to = "parameter",
               values_to = "n")

GLODAP_stats_long  %>%  write_csv(
  here::here(
    "data/GLODAPv2_2020/_summarized_data_files",
    "GLODAPv2.2020_stats.csv"
  )
)

##

GLODAP_obs_grid_clean %>% 
  write_csv(
    here::here(
      "data/GLODAPv2_2020/_summarized_data_files",
      "GLODAPv2.2020_clean_obs_grid.csv"
    )
  )

##

GLODAP <- GLODAP %>% 
  select(year, date, era, basin, lat, lon, cruise,
         bottomdepth, depth, tem, sal, gamma,
         tco2, talk, phosphate,
         oxygen, aou, nitrate, silicate)

GLODAP  %>%  write_csv(
  here::here(
    "data/GLODAPv2_2020/_summarized_data_files",
    "GLODAPv2.2020_clean.csv"
  )
)

# GLODAP  %>%
#   filter(year == 2007) %>%
#   select(date, lat, lon, depth, tem, sal, tco2:silicate) %>%
#   write_csv(
#     here::here(
#       "data/GLODAPv2_2020/_summarized_data_files",
#       "GLODAPv2.2020_clean_2007.csv"
#     )
#   )

```


# Overview plots

## Cleaning stats

Number of observations at various steps of data cleaning.

```{r GLODAP_cleaning_stats}

GLODAP_stats_long <- GLODAP_stats_long %>%
  mutate(parameter = fct_reorder(parameter, n))

GLODAP_stats_long %>% 
  ggplot(aes(parameter, n/1000)) +
  geom_col() +
  coord_flip() +
  labs(y = "n (1000)") +
  theme(axis.title.y = element_blank())

rm(GLODAP_stats_long)

```


## Assign coarse spatial grid

For the following plots, the cleaned data set was re-opened and observations were gridded spatially to intervals of:  

- 5° x 5°

```{r grid_spatially_5x5}

GLODAP <- GLODAP %>% 
  mutate(lat_grid = cut(lat, seq(-90, 90, 5), seq(-87.5, 87.5, 5)),
         lat_grid = as.numeric(as.character(lat_grid)),
         lon_grid = cut(lon, seq(20, 380, 5), seq(22.5, 377.5, 5)),
         lon_grid = as.numeric(as.character(lon_grid)))

```


## Histogram Zonal coverage

```{r coverage_histogram_zonal}

GLODAP_histogram_lat <- GLODAP %>%
  group_by(era, lat_grid, basin) %>%
  tally() %>%
  ungroup()

GLODAP_histogram_lat %>%
  ggplot(aes(lat_grid, n, fill = era)) +
  geom_col() +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap( ~ basin) +
  coord_flip(expand = 0) +
  theme(legend.position = "top",
        legend.title = element_blank())

rm(GLODAP_histogram_lat)

```

## Histogram temporal coverage

```{r coverage_histogram_temporal}

GLODAP_histogram_year <- GLODAP %>%
  group_by(year, basin) %>%
  tally() %>%
  ungroup()

era_median_year <- GLODAP %>%
  group_by(era) %>%
  summarise(t_ref = median(year)) %>%
  ungroup()

GLODAP_histogram_year %>%
  ggplot() +
  geom_vline(xintercept = c(
    parameters$year_JGOFS_end + 0.5,
    parameters$year_GOSHIP_end + 0.5
  )) +
  geom_col(aes(year, n, fill = basin)) +
  geom_point(
    data = era_median_year,
    aes(t_ref, 0, shape = "Median year"),
    size = 2,
    fill = "white"
  ) +
  scale_fill_brewer(palette = "Dark2") +
  scale_shape_manual(values = 24, name = "") +
  scale_y_continuous() +
  coord_cartesian(expand = 0) +
  theme(
    legend.position = "top",
    legend.direction = "vertical",
    legend.title = element_blank(),
    axis.title.x = element_blank()
  )

rm(GLODAP_histogram_year,
   era_median_year)

```

## Zonal temporal coverage (Hovmoeller)

```{r coverage_hovmoeller, fig.asp=1}

GLODAP_hovmoeller_year <- GLODAP %>%
  group_by(year, lat_grid, basin) %>%
  tally() %>%
  ungroup()

GLODAP_hovmoeller_year %>%
  ggplot(aes(year, lat_grid, fill = log10(n))) +
  geom_tile() +
  geom_vline(xintercept = c(1999.5, 2012.5)) +
  scale_fill_viridis_c(option = "magma", direction = -1) +
  facet_wrap( ~ basin, ncol = 1) +
  theme(legend.position = "top",
        axis.title.x = element_blank())

rm(GLODAP_hovmoeller_year)

```

## Coverage maps by era

### Subsetting process

```{r coverage_maps_era_subsetting, fig.asp=1.5}

GLODAP_obs_grid <- GLODAP_obs_grid %>%
  mutate(era = factor(era, c("JGOFS_WOCE", "GO_SHIP", "new_era")),
         cleaning_level = factor(
           cleaning_level,
           unique(GLODAP_obs_grid$cleaning_level)
         ))

ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat),
              fill = "grey80") +
  geom_raster(data = GLODAP_obs_grid %>%
                filter(cleaning_level == "all") %>% 
                select(-cleaning_level),
              aes(lon, lat, fill = "all")) +
  geom_raster(data = GLODAP_obs_grid %>%
                filter(cleaning_level != "all"),
              aes(lon, lat, fill = "subset")) +
  coord_quickmap(expand = FALSE) +
  scale_fill_manual(values = c("grey", "red"), name = "") +
  facet_grid(cleaning_level ~ era) +
  theme(legend.position = "top",
        axis.title = element_blank())

```

### eMLR variables

Grey pixels refer to sampling locations filtered for availability of tco2 data only.

```{r coverage_maps_era_clean, fig.asp=1.5}

GLODAP <- GLODAP %>% 
  mutate(era = factor(era, c("JGOFS_WOCE", "GO_SHIP", "new_era")))

GLODAP_tco2_grid <- GLODAP %>%
  count(lat, lon, era)

GLODAP %>%
  ggplot(aes(lon, lat)) +
  geom_raster(data = landmask,
              aes(lon, lat), fill = "grey80") +
  geom_raster(data = GLODAP_tco2_grid, aes(lon, lat), fill = "grey80") +
  geom_bin2d(binwidth = c(1,1)) +
  scale_fill_viridis_c(option = "magma", direction = -1, trans = "log10",
                       name = "log10(eMLR_variables)") +
  coord_quickmap(expand = FALSE) +
  facet_wrap(~era, ncol = 1) +
  theme(legend.position = "top",
        axis.title = element_blank())

```
