---
title: "GLODAP"
author: "Jens Daniel M체ller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(patchwork)
```


```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

# Read master file

- Data source: `GLODAPv2.2020_Merged_Master_File.csv` from [glodap.info](https://www.glodap.info/)

```{r read_GLODAPv2_2020_merged_master_file, eval=FALSE}

GLODAP <- read_csv(here::here("data/GLODAPv2_2020/Merged_data_product",
                              "GLODAPv2.2020_Merged_Master_File.csv"),
                   na = "-9999",
                   col_types = cols(.default = col_double()))

# relevant columns
GLODAP <- GLODAP %>% 
  select(cruise:talkqc)

GLODAP <- GLODAP %>% 
  mutate(date = ymd(paste(year, month, day))) %>% 
         #decade = as.factor(floor(year / 10) * 10)) %>% 
  relocate(date)

GLODAP <- GLODAP %>% 
  select(-c(month:minute, 
            maxsampdepth, pressure,
            theta, sigma0:gamma,
            nitrate:nitritef))

```

# Clean data

- subsetted rows with 
  - paired observations of all relevant parameters
  - flag f: 2
  - flag qc: 1
  
- calculate summary statistics during cleaning process

```{r clean_GLODAP_master_file, eval=FALSE}

GLODAP_stats <- GLODAP %>% 
  summarise(total = n())

GLODAP <- GLODAP %>% 
  filter(!is.na(tco2))

GLODAP <- GLODAP %>% 
  filter(tco2f == "2") %>% 
  select(-tco2f)

GLODAP <- GLODAP %>% 
  filter(tco2qc == "1") %>% 
  select(-tco2qc)

GLODAP_stats_temp <- GLODAP %>% 
  summarise(tco2 = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP <- GLODAP %>% 
  filter(!is.na(talk))

GLODAP <- GLODAP %>% 
  filter(talkf == "2") %>% 
  select(-talkf)

GLODAP <- GLODAP %>% 
  filter(talkqc == "1") %>% 
  select(-talkqc)

##

GLODAP <- GLODAP %>% 
  filter(!is.na(phosphate))

GLODAP <- GLODAP %>% 
  filter(phosphatef == "2") %>% 
  select(-phosphatef)

GLODAP <- GLODAP %>% 
  filter(phosphateqc == "1") %>% 
  select(-phosphateqc)

GLODAP_stats_temp <- GLODAP %>% 
  summarise(C_star_variables = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP <- GLODAP %>% 
  filter(!is.na(temperature))

##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(salinity))

GLODAP <- GLODAP %>% 
  filter(salinityf == "2") %>% 
  select(-salinityf)

GLODAP <- GLODAP %>% 
  filter(salinityqc == "1") %>% 
  select(-salinityqc)

##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(silicate))

GLODAP <- GLODAP %>% 
  filter(silicatef == "2") %>% 
  select(-silicatef)

GLODAP <- GLODAP %>% 
  filter(silicateqc == "1") %>% 
  select(-silicateqc)

##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(oxygen))

GLODAP <- GLODAP %>% 
  filter(oxygenf == "2") %>% 
  select(-oxygenf)

GLODAP <- GLODAP %>% 
  filter(oxygenqc == "1") %>% 
  select(-oxygenqc)

##

GLODAP <- GLODAP %>% 
  filter(!is.na(aou))

GLODAP <- GLODAP %>% 
  filter(aouf == "2") %>% 
  select(-aouf)

GLODAP_stats_temp <- GLODAP %>% 
  summarise(eMLR_variables = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)


GLODAP_stats_long <- GLODAP_stats %>% 
  pivot_longer(1:4, names_to = "parameter", values_to = "n")

GLODAP_stats_long  %>%  write_csv(here::here("data/GLODAPv2_2020/_summarized_data_files",
                                             "GLODAPv2.2020_stats.csv"))

rm(GLODAP_stats_long, GLODAP_stats)

```

# Assign reference eras

Samples were assigned to following eras:

`JGOFS/WOCE` 1981 - 1999

`GO-SHIP` 2000 - 2012
 
`new_era` 2013 - now

```{r era_assignement, eval=FALSE}

GLODAP <- GLODAP %>%
  filter(year >= 1981) %>% 
  mutate(era = "JGOFS/WOCE",
         era = if_else(year >= 2000, "GO-SHIP", era),
         era = if_else(year >= 2013, "new_era", era))

```
 
```{r write_clean_data_file, eval=FALSE}

GLODAP  %>%  write_csv(here::here("data/GLODAPv2_2020/_summarized_data_files",
                                  "GLODAPv2.2020_clean.csv"))

```


# Overview plots

Open clean data file.

```{r reopen_GLODAP}

GLODAP <- read_csv(here::here("data/GLODAPv2_2020/_summarized_data_files",
                              "GLODAPv2.2020_clean.csv"))

```

## Assign spatial grid

For the following plots, observations were gridded spatially to 5째x5째 intervals.

```{r grid_spatially}

GLODAP <- GLODAP %>% 
  mutate(lat_grid = cut(latitude, seq(-90, 90, 5), seq(-87.5, 87.5, 5)),
         lat_grid = as.numeric(as.character(lat_grid)),
         lon_grid = cut(longitude, seq(-180, 180, 5), seq(-177.5, 177.5, 5)),
         lon_grid = as.numeric(as.character(lon_grid))) %>% 
  arrange(date)

```


## Cleaning stats

```{r GLODAP_stats_barplot}

GLODAP_stats_long <-  read_csv(here::here("data/GLODAPv2_2020/_summarized_data_files",
                                             "GLODAPv2.2020_stats.csv"))

GLODAP_stats_long <- GLODAP_stats_long %>%
  mutate(parameter = fct_reorder(parameter, n))

GLODAP_stats_long %>% 
  ggplot(aes(parameter, n/1000))+
  geom_col()+
  coord_flip()+
  labs(y="n (1000)")+
  theme(axis.title.y = element_blank())

rm(GLODAP_stats_long)

```

## Histogram latitudes

```{r histogramm_latitude_observations}

GLODAP_histogram_lat <- GLODAP %>% 
  group_by(era, lat_grid) %>% 
  tally() %>% 
  ungroup()

GLODAP_histogram_lat %>% 
  ggplot(aes(lat_grid, n, fill=era))+
  geom_col()+
  scale_x_continuous(breaks = seq(-87.5,90,10))+
  scale_fill_viridis_d()+
  coord_flip(expand = 0)

rm(GLODAP_histogram_lat)

```

## Histogram year

```{r histogramm_year_observations}

GLODAP_histogram_year <- GLODAP %>% 
  group_by(era, year) %>% 
  tally() %>% 
  ungroup()

era_median_year <- GLODAP %>% 
  group_by(era) %>% 
  summarise(t_ref = median(year)) %>% 
  ungroup()

GLODAP_histogram_year %>% 
  ggplot(aes(year, n, fill=era))+
  geom_col()+
  geom_vline(data = era_median_year, aes(xintercept = t_ref))+
  coord_cartesian(expand = 0)


rm(GLODAP_histogram_year,
   era_median_year)

```

## Hovmoeller (lat vs year)

```{r coverage_hovmoeller}

GLODAP_hovmoeller_year <- GLODAP %>% 
  group_by(year, lat_grid) %>% 
  tally() %>% 
  ungroup()

GLODAP_hovmoeller_year %>% 
  ggplot(aes(year, lat_grid, fill=log10(n)))+
  geom_tile()+
  geom_vline(xintercept = c(1999.5, 2012.5))+
  scale_fill_viridis_c(option = "magma", direction = -1)

rm(GLODAP_hovmoeller_year)

```

## Maps by era

```{r coverage_maps_era, fig.asp=1.5}

mapWorld <- borders("world", colour="gray60", fill="gray60")

GLODAP_map_era <- GLODAP %>% 
  group_by(era, lat_grid, lon_grid) %>% 
  tally() %>% 
  ungroup()

GLODAP_map_era <- GLODAP_map_era %>% 
  mutate(era = factor(era, c("JGOFS/WOCE", "GO-SHIP", "new_era")))

GLODAP_map_era %>%
  ggplot(aes(lon_grid, lat_grid, fill=log10(n)))+
  mapWorld+
  geom_raster()+
  scale_fill_viridis_c(option = "magma", direction = -1)+
  scale_x_continuous(breaks = seq(-180, 180, 30))+
  scale_y_continuous(breaks = seq(-90, 90, 30))+
  coord_quickmap(expand = FALSE)+
  facet_wrap(~era, ncol=1)

rm(GLODAP_map_era,
   mapWorld)

```

# Individual cruise sections

```{r individual_cruises_clean_triplicates, eval = FALSE}

cruises <- GLODAP %>% 
  group_by(cruise) %>% 
  summarise(date_mean = mean(date, na.rm = TRUE),
            n = n()) %>% 
  ungroup() %>% 
  arrange(date_mean)

GLODAP <- full_join(GLODAP, cruises)

mapWorld <- borders("world", colour="gray60", fill="gray60")

n <- 0
for (i_cruise in unique(cruises$cruise)) {

#i_cruise <- unique(cruises$cruise)[1]
n <- n+1
print(n)  
  
GLODAP_cruise <- GLODAP %>%
  filter(cruise == i_cruise) %>% 
  arrange(date)

cruises_cruise <- cruises %>%
  filter(cruise == i_cruise)
  
map <- GLODAP_cruise %>%
  ggplot(aes(longitude, latitude))+
  mapWorld+
  geom_point(aes(col=date))+
  geom_path()+
  coord_quickmap(expand = FALSE)+
  scale_color_viridis_c(trans = "date")+
  labs(title = paste("Mean date:", cruises_cruise$date_mean,
                     "| cruise:", cruises_cruise$cruise,
                     "| n(samples):", cruises_cruise$n))


lon_section <- GLODAP_cruise %>%
  ggplot(aes(longitude, depth))+
  scale_y_reverse()+
  scale_color_viridis_c()

lon_tco2 <- lon_section+
  geom_point(aes(col=tco2))

lon_talk <- lon_section+
  geom_point(aes(col=talk))

lon_phosphate <- lon_section+
  geom_point(aes(col=phosphate))


lat_section <- GLODAP_cruise %>%
  ggplot(aes(latitude, depth))+
  scale_y_reverse()+
  scale_color_viridis_c()

lat_tco2 <- lat_section+
  geom_point(aes(col=tco2))

lat_talk <- lat_section+
  geom_point(aes(col=talk))

lat_phosphate <- lat_section+
  geom_point(aes(col=phosphate))

map /
((lat_tco2 / lat_talk / lat_phosphate) |
(lon_tco2 / lon_talk / lon_phosphate))

ggsave(here::here("output/plots/data/all_cruises_clean",
                  paste("GLODAP_cruise_date",
                        cruises_cruise$date_mean,
                        "n",
                        cruises_cruise$n,
                        "cruise",
                        cruises_cruise$cruise,
                        ".png",
                        sep = "_")),
                  width = 9, height = 9)

rm(map,
   lon_section, lat_section,
   lat_tco2, lat_talk, lat_phosphate, lon_tco2, lon_talk, lon_phosphate,
   GLODAP_cruise, cruises_cruise)

}


# library("rnaturalearth")
# library("rnaturalearthdata")
# library("sf")
# 
# world <- ne_countries(scale = "small", returnclass = "sf")
# class(world)
# 
# GLODAP_map <- GLODAP %>% 
#   group_by(lat_grid, lon_grid) %>% 
#   tally() %>% 
#   ungroup()
# 
# ggplot() +
#   geom_raster(data = GLODAP_map, aes(lon_grid, lat_grid))+
#   geom_sf(data = world)+
#   coord_sf(crs = "+proj=robin +lat_0=0 +lon_0=0 +x0=0 +y0=0")


# https://gist.github.com/clauswilke/783e1a8ee3233775c9c3b8bfe531e28a
# https://twitter.com/clauswilke/status/1066024436208406529
# https://www.r-spatial.org/r/2018/10/25/ggplot2-sf.html
```

# Open tasks

- remove marginal seas and data north of 65째N
- move A16 cruises from new_era to GO-SHIP era

# Questions

- exclude coastal area in general?
